---
title: "28 Ind. Peak Quant"
author: "Briana Mittleman"
date: "8/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I know have 28 individuals sequences on 2 lanes. I have combined these and used the peak coverage pipeline to call and clean peaks. I will use this analysis to explore the library sizes and coverage at these peaks.  


```{r}
library(tidyverse)
library(workflowr)
library(cowplot)
library(reshape2)
library(devtools)
```

##Reads and Mapping Stats:  

```{r}
map_stats=read.csv("../data/comb_map_stats.csv", header=T)
map_stats$line=as.factor(map_stats$line)
map_stats$batch=as.factor(map_stats$batch)
```

The number of reads for each library and the number of mapped reads.  

```{r}
read_plot=ggplot(map_stats, aes(x=line, y=comb_reads, fill=fraction))+ geom_bar(stat="identity", position="dodge") +labs(y="Reads", title="Reads by line and fraction") 
map_plot=ggplot(map_stats, aes(x=line, y=comb_mapped, fill=fraction))+ geom_bar(stat="identity", position="dodge") +labs(y="Mapped Reads", title="Mapped reads by line and fraction") + geom_hline(yintercept=10000000) + annotate("text",label="10 million mapped reads", y=9000000, x=10)

plot_grid(read_plot, map_plot)
```


The percent of reads that map per line are pretty uniform accross libraries. The mean is 72%. 

```{r}
ggplot(map_stats, aes(x=line, y=comb_prop_mapped, fill=fraction))+ geom_bar(stat="identity", position="dodge") +labs(y="Mapped Percent", title="Percent of reads mapping by line and fraction") 

mean(map_stats$comb_prop_mapped)
```

##Clean peak exploration  

```{r}
peak_quant=read.table(file = "../data/clean_peaks/APAquant.fc.cleanpeaks.fc", header=T)
```

Fix the names 
```{r}
file_names=colnames(peak_quant)[7:62]
file_names_split=lapply(file_names, function(x)strsplit(x,".", fixed=T))
libraries=c()
for (i in file_names_split){
  unlist_i=unlist(i)
  libraries=c(libraries, paste(unlist_i[10], unlist_i[11], sep="-"))
}
colnames(peak_quant)=c(colnames(peak_quant)[1:6], libraries)  
```


Explore the peaks before quantifications:  

```{r}
#length of peaks
plot(sort(peak_quant$Length,decreasing = T), main="Peak Lengths", ylab="Peak Length", xlab="Peak index")


#mean cov of peaks

peak_cov=peak_quant %>% select(contains("-"))

peak_mean=apply(peak_cov,1,mean)
peak_var=apply(peak_cov, 1, var)

plot(log10(sort(peak_mean,decreasing = T)))
plot(peak_var)


plot(log10(peak_var)~log10(peak_mean))


```

Plot the coverage vs the length:  

```{r}
plot(peak_mean~peak_quant$Length)
```

##Clustering:  


```{r}
cor_function=function(data){
  corr_matrix= matrix(0,ncol(data),ncol(data))
  for (i in seq(1,ncol(data))){
    for (j in seq(1,ncol(data))){
      x=cor.test(data[,i],  data[,j], method='pearson')
      cor_ij=as.numeric(x$estimate)
      corr_matrix[i,j]=cor_ij
    }
  }
  return(corr_matrix)
}

count_cor=cor_function(peak_cov)
rownames(count_cor)=libraries
colnames(count_cor)=libraries


```


```{r}
melted_count_corr=melt(count_cor)
ggheatmap=ggplot(data = melted_count_corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  labs(title="Correlation Heatplot")


ggheatmap
```


CLustering:  

```{r}
pca_peak= prcomp(peak_cov,center = TRUE,scale. = TRUE)
summary(pca_peak)
```
```{r}

pc_df=as.data.frame(pca_peak$rotation) %>% rownames_to_column(var="lib") %>% mutate(fraction=ifelse(grepl("T", lib), "total", "nuclear"))

ggplot(pc_df, aes(x=PC1, y=PC2, col=fraction)) + geom_point() + labs(x="PC1: Prop explained 0.7763", y="PC2: Prop explained 0.1330", title="Raw PAS qunatification data")
```


I now want to explore what the first PC is representing. Some ideas are:  

* batch  

* sequencing depth  

* mapped reads  


All of this info is in the map stats.  

```{r}
pc_df=as.data.frame(pca_peak$rotation) %>% rownames_to_column(var="lib") %>% mutate(fraction=ifelse(grepl("T", lib), "total", "nuclear")) %>% mutate(reads=map_stats$comb_reads) %>% mutate(batch=map_stats$batch) %>% mutate(mapped=map_stats$comb_mapped)


batch_gg= ggplot(pc_df, aes(x=PC1, y=PC2, col=batch)) + geom_point() + labs(x="PC1: Prop explained 0.7763", y="PC2: Prop explained 0.1330", title="Batch")

frac_gg= ggplot(pc_df, aes(x=PC1, y=PC2, col=fraction)) + geom_point() + labs(x="PC1: Prop explained 0.7763", y="PC2: Prop explained 0.1330", title="Fraction")

reads_gg= ggplot(pc_df, aes(x=PC1, y=PC2, col=reads)) + geom_point() + labs(x="PC1: Prop explained 0.7763", y="PC2: Prop explained 0.1330", title="Reads")


mapped_gg= ggplot(pc_df, aes(x=PC1, y=PC2, col=mapped)) + geom_point() + labs(x="PC1: Prop explained 0.7763", y="PC2: Prop explained 0.1330", title="Mapped Reads")

plot_grid(frac_gg,batch_gg,reads_gg,mapped_gg)


```

Proportion of reads mapping to peaks. This may be in the feature counts summary.  


