---
title: "apaQTLs by gene count percentile"
author: "Briana Mittleman"
date: "2/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will use this analysis to look at the number of apaQTL genes by the percentile of the counts for the gene. This may help us know if we want to sequence the libraries deeper.  



```{r}
library(workflowr)
library(tidyverse)
library(cowplot)
library(reshape2)
```

##QTL genes  
First, upload the QTL gene.  
```{r}
nucQTLs=read.table("../data/ApaQTLs/NuclearapaQTLs.GeneLocAnno.noMP.5perc.10FDR.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval", "bh"))

nucQTLsGenes= nucQTLs%>%separate(pid, into=c("chr", "start", "end", "id"), sep=":") %>% separate(id, into=c("gene", "strand", "peak"), sep="_") %>% select(gene) %>% unique()


totQTLs=read.table("../data/ApaQTLs/TotalapaQTLs.GeneLocAnno.noMP.5perc.10FDR.txt",stringsAsFactors = F, col.names = c("pid", "nvar", "shape1", "shape2", "dummy", "sid", "dist", "npval", "slope", "ppval", "bpval", "bh"))

totQTLsGenes= totQTLs%>%separate(pid, into=c("chr", "start", "end", "id"), sep=":") %>% separate(id, into=c("gene", "strand", "peak"), sep="_") %>% select(gene) %>% unique()
  
```

How many of the genes overlap:
```{r}
QTLgene_both=totQTLsGenes %>% inner_join(nucQTLsGenes, by="gene")
nrow(QTLgene_both) 
```
This means 141 genes have a QTL for both. It does not tell me if the QTL is the same.


##Gene Counts  

I can get this from the prefiltered peak counts. /project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.fc  

/project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov_noMP_GeneLocAnno/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.fc  


I can pull these files in, group them by gene and get summaries. 


###Total  
```{r}
totPeakCounts=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(-peak, -chr,-start, -end, -strand, -gene)
totPeakCountsGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene)
#sum across ind.
totPeakCounts_Sum=rowSums(totPeakCounts)
totPeakCountsGeneSum=as.data.frame(cbind(totPeakCountsGene,totPeakCounts_Sum)) %>% group_by(gene) %>% summarise(TotalCount=sum(totPeakCounts_Sum)) %>% mutate(Percentile = percent_rank(TotalCount)) 

```


Subset by percetile:  

```{r}
totalCount_10= totPeakCountsGeneSum %>% filter(Percentile<.1)
totalCount_20= totPeakCountsGeneSum %>% filter(Percentile<.2, Percentile>.1)
totalCount_30= totPeakCountsGeneSum %>% filter(Percentile<.3, Percentile>.2)
totalCount_40= totPeakCountsGeneSum %>% filter(Percentile<.4, Percentile>.3)
totalCount_50= totPeakCountsGeneSum %>% filter(Percentile<.5, Percentile>.4)
totalCount_60= totPeakCountsGeneSum %>% filter(Percentile<.6, Percentile>.5)
totalCount_70= totPeakCountsGeneSum %>% filter(Percentile<.7, Percentile>.6)
totalCount_80= totPeakCountsGeneSum %>% filter(Percentile<.8, Percentile>.7)
totalCount_90= totPeakCountsGeneSum %>% filter(Percentile<.9, Percentile>.8)
totalCount_100= totPeakCountsGeneSum %>% filter(Percentile<1, Percentile>.9)

```


###Nuclear

```{r}
nucPeakCounts=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(-peak, -chr,-start, -end, -strand, -gene)
nucPeakCountsGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene)
#sum across ind.
nucPeakCounts_Sum=rowSums(nucPeakCounts)
nucPeakCountsGeneSum=as.data.frame(cbind(nucPeakCountsGene,nucPeakCounts_Sum)) %>% group_by(gene) %>% summarise(NuclearCount=sum(nucPeakCounts_Sum)) %>% mutate(Percentile = percent_rank(NuclearCount)) 

```

```{r}
nuclearCount_10= nucPeakCountsGeneSum %>% filter(Percentile<.1)
nuclearCount_20= nucPeakCountsGeneSum %>% filter(Percentile<.2, Percentile>.1)
nuclearCount_30= nucPeakCountsGeneSum %>% filter(Percentile<.3, Percentile>.2)
nuclearCount_40= nucPeakCountsGeneSum %>% filter(Percentile<.4, Percentile>.3)
nuclearCount_50= nucPeakCountsGeneSum %>% filter(Percentile<.5, Percentile>.4)
nuclearCount_60= nucPeakCountsGeneSum %>% filter(Percentile<.6, Percentile>.5)
nuclearCount_70= nucPeakCountsGeneSum %>% filter(Percentile<.7, Percentile>.6)
nuclearCount_80= nucPeakCountsGeneSum %>% filter(Percentile<.8, Percentile>.7)
nuclearCount_90= nucPeakCountsGeneSum %>% filter(Percentile<.9, Percentile>.8)
nuclearCount_100= nucPeakCountsGeneSum %>% filter(Percentile<1, Percentile>.9)

```


##QTL genes  

I can get the percentile for each QTL gene.  

```{r}
totQTLGene_Perc= totQTLsGenes %>% inner_join(totPeakCountsGeneSum, by="gene") %>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n())
nucQTLGene_Perc= nucQTLsGenes %>% inner_join(nucPeakCountsGeneSum, by="gene")%>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n())

bothPerc=totQTLGene_Perc %>% full_join(nucQTLGene_Perc,  by = c("roundPerc"))
bothPerc$Ngenes.x= bothPerc$Ngenes.x %>% replace_na(0)
colnames(bothPerc)= c("Percentile", "Total", "Nuclear")


bothPerc_melt= melt(bothPerc, id.vars = "Percentile") 
colnames(bothPerc_melt) =c("Percentile", "Fraction", "Genes")
```

Plot: 


```{r}
ggplot(bothPerc_melt, aes(x=Percentile, y= Genes, by=Fraction, fill=Fraction)) + geom_bar(stat="identity", position = "dodge") + theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  
```

To make is a percente of genes in the category I will divide each number of genes category by the number of total QTL genes. For total this is  291 and for nuclear it is 496

```{r}
totQTLGene_Perc_prop= totQTLsGenes %>% inner_join(totPeakCountsGeneSum, by="gene") %>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n()) %>% mutate(PercGenes=Ngenes/291)
nucQTLGene_Perc_prop= nucQTLsGenes %>% inner_join(nucPeakCountsGeneSum, by="gene")%>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n()) %>%mutate(PercGenes=Ngenes/496)


bothPercProp=totQTLGene_Perc_prop %>% full_join(nucQTLGene_Perc_prop,  by = c("roundPerc")) %>% select(roundPerc, starts_with("perc"))
bothPercProp$PercGenes.x= bothPercProp$PercGenes.x %>% replace_na(0)
colnames(bothPercProp)= c("Percentile", "Total", "Nuclear")


bothPercPrp_melt= melt(bothPercProp, id.vars = "Percentile") 
colnames(bothPercPrp_melt) =c("Percentile", "Fraction", "GenesProp")
```

```{r}
QTLSbyCountPerc=ggplot(bothPercPrp_melt, aes(x=Percentile, y=GenesProp, fill=Fraction)) +geom_bar(stat="identity", position = "dodge")+labs(title="Proportion of QTL genes by Read count percentile",y="Proportion of QTLs", x="Read Cound percentile") +  theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  + facet_grid(~Fraction)
QTLSbyCountPerc

ggsave(QTLSbyCountPerc, file="../output/plots/QTLSbyCountPerc.png")
```


Filter to only look at genes with at least 2 peaks and build the percentiles off these. To do this I will filter the unfiltered peak counts by genes with two peaks in the 5% peaks used in the QTL analysis.    

Pull in filtered peaks:  
```{r}
Peaks=read.table("../data/PeakUsage_noMP_GeneLocAnno/Filtered_APApeaks_merged_allchrom_noMP.sort.named.noCHR_geneLocParsed.5percCov.bed", stringsAsFactors = F, col.names = c("chr", 'start', 'end', 'id', 'score', 'strand')) 
Genes2Peaks= Peaks %>% separate(id, into=c("gene", "peak"), sep=":") %>% group_by(gene) %>% summarise(nPeak=n()) %>% filter(nPeak>=2) %>% select(gene)
```


Now filter the total and nuclear peaks that are in these genes before making the percentile plots.  

Total
```{r}
totPeakCounts_FiltGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% semi_join(Genes2Peaks, by="gene")%>% select(-peak, -chr,-start, -end, -strand, -gene)

totPeakCountsFiltGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% semi_join(Genes2Peaks, by="gene") %>% select(gene)
#sum across ind.
totPeakCounts_FiltSum=rowSums(totPeakCounts_FiltGene)
totPeakCountsFiltGeneSum=as.data.frame(cbind(totPeakCountsFiltGene,totPeakCounts_FiltSum)) %>% group_by(gene) %>% summarise(TotalCount=sum(totPeakCounts_FiltSum)) %>% mutate(Percentile = percent_rank(TotalCount)) 
```

Nuclear  

```{r}
nucPeakCounts_FiltGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% semi_join(Genes2Peaks, by="gene")%>% select(-peak, -chr,-start, -end, -strand, -gene)

nucPeakCountsFiltGene=read.table("../data/AllPeak_counts/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.fc", header = T, stringsAsFactors = F) %>% select(-Chr, -Start, -End, -Strand, -Length) %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% semi_join(Genes2Peaks, by="gene") %>% select(gene)
#sum across ind.
nucPeakCounts_FiltSum=rowSums(nucPeakCounts_FiltGene)
nucPeakCountsFiltGeneSum=as.data.frame(cbind(nucPeakCountsFiltGene,nucPeakCounts_FiltSum)) %>% group_by(gene) %>% summarise(NuclearCount=sum(nucPeakCounts_FiltSum)) %>% mutate(Percentile = percent_rank(NuclearCount)) 
```


I can get the percentile for each QTL gene.  

```{r}
totQTLFiltGene_Perc= totQTLsGenes %>% inner_join(totPeakCountsFiltGeneSum, by="gene") %>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n())
nucQTLFiltGene_Perc= nucQTLsGenes %>% inner_join(nucPeakCountsFiltGeneSum, by="gene")%>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n())

bothPerc_filt=totQTLFiltGene_Perc %>% full_join(nucQTLFiltGene_Perc,  by = c("roundPerc"))
bothPerc_filt$Ngenes.x= bothPerc_filt$Ngenes.x %>% replace_na(0)
colnames(bothPerc_filt)= c("Percentile", "Total", "Nuclear")


bothPercfilt_melt= melt(bothPerc_filt, id.vars = "Percentile") 
colnames(bothPercfilt_melt) =c("Percentile", "Fraction", "Genes")
```

Plot: 


```{r}
ggplot(bothPercfilt_melt, aes(x=Percentile, y= Genes, by=Fraction, fill=Fraction)) + geom_bar(stat="identity", position = "dodge") + theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  
```

Do it by percent  

```{r}
totQTLFiltGene_Perc_prop= totQTLsGenes %>% inner_join(totPeakCountsFiltGeneSum, by="gene") %>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n()) %>% mutate(PercGenes=Ngenes/291)
nucQTLFiltGene_Perc_prop= nucQTLsGenes %>% inner_join(nucPeakCountsFiltGeneSum, by="gene")%>% mutate(roundPerc=round(Percentile, digits=1)) %>% group_by(roundPerc) %>% summarise(Ngenes=n()) %>%mutate(PercGenes=Ngenes/496)


bothPercFiltProp=totQTLFiltGene_Perc_prop %>% full_join(nucQTLFiltGene_Perc_prop,  by = c("roundPerc")) %>% select(roundPerc, starts_with("perc"))
bothPercFiltProp$PercGenes.x= bothPercFiltProp$PercGenes.x %>% replace_na(0)
colnames(bothPercFiltProp)= c("Percentile", "Total", "Nuclear")


bothPercPrpFilt_melt= melt(bothPercFiltProp, id.vars = "Percentile") 
colnames(bothPercPrpFilt_melt) =c("Percentile", "Fraction", "GenesProp")
```

```{r}
QTLSbyCountPerc_filt=ggplot(bothPercPrpFilt_melt, aes(x=Percentile, y=GenesProp, fill=Fraction)) +geom_bar(stat="identity", position = "dodge")+labs(title="Proportion of QTL genes by Read count percentile\n Genes with 2 peaks in QTL analysis",y="Proportion of QTLs", x="Read Cound percentile") +  theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  + facet_grid(~Fraction)
QTLSbyCountPerc_filt

ggsave(QTLSbyCountPerc_filt, file="../output/plots/QTLSbyCountPerc_Genes2Peaks.png")
```
