---
title: "Characterize Total ApaQTLs"
author: "Briana Mittleman"
date: "10/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This analysis will be used to characterize the total ApaQTLs. I will run the analysis on the total APAqtls in this analysis and will then run a similar analysis on the nuclear APAqtls in another analysis. I would like to study:  


* Distance metrics:  
    + distance from snp to TSS of gene  
    + Distance from snp to peak  

* Expression metrics: 
    + expression of genes with significant QTLs vs other genes  (by rna seq)
    + expression of genes with significant QTLs vs other genes  (peak coverage)  

* Chrom HMM metrics:  
    + look at the chrom HMM interval for the significant QTLs  


 




##Upload Libraries and Data:  

Library
```{r}
library(workflowr)
library(reshape2)
library(tidyverse)
library(VennDiagram)
library(data.table)
library(cowplot)
```


Permuted Results from APA:

I will add a column to this dataframe that will tell me if the association is significant at 10% FDR.  This will help me plot based on significance later in the analysis. I am also going to seperate the PID into relevant pieces.  

```{r}
totalAPA=read.table("../data/perm_QTL_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Total_transcript_permResBH.txt", stringsAsFactors = F, header=T)  %>% mutate(sig=ifelse(-log10(bh)>=1, 1,0 )) %>%  separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak"))

totalAPA$sig=as.factor(totalAPA$sig)


print(names(totalAPA))
```



##Distance Metrics  

### Distance from snp to TSS  

I ran the QTL analysis based on the starting position of the gene. 


```{r}
ggplot(totalAPA, aes(x=dist, fill=sig, by=sig)) + geom_density(alpha=.5)  +  labs(title="Distance from snp to TSS", x="Base Pairs") + scale_fill_discrete(guide = guide_legend(title = "Significant QTL"))
```

It looks like most of the signifcant values are 100,000 bases. This makes sense.  I can zoom in on this portion.  

```{r}
ggplot(totalAPA, aes(x=dist, fill=sig, by=sig)) + geom_density(alpha=.5)+coord_cartesian(xlim = c(-150000, 150000))
```

### Distance from snp to peak  

To perform this analysis I need to recover the peak positions.  

The peak file I used for the QTL analysis is: /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed.bed  

```{r}
peaks=read.table("../data/PeaksUsed/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed.bed", col.names = c("chr", "peakStart", "peakEnd", "PeakNum", "PeakScore", "Strand", "Gene")) %>% mutate(peak=paste("peak", PeakNum,sep="")) %>% mutate(PeakCenter=peakStart+ (peakEnd- peakStart))
```

I want to join the peak start to the totalAPA file but the peak column. I will then create a column that is snppos-peakcenter.

```{r}
totalAPA_peakdist= totalAPA %>%  inner_join(peaks, by="peak") %>%  separate(sid, into=c("snpCHR", "snpLOC"), by=":")
totalAPA_peakdist$snpLOC= as.numeric(totalAPA_peakdist$snpLOC)

totalAPA_peakdist= totalAPA_peakdist %>%  mutate(DisttoPeak= snpLOC-PeakCenter)
```

Plot this by significance.  
```{r}
ggplot(totalAPA_peakdist, aes(x=DisttoPeak, fill=sig, by=sig)) + geom_density(alpha=.5)  +  labs(title="Distance from snp peak", x="log10 absolute value Distance to Peak") + scale_fill_discrete(guide = guide_legend(title = "Significant QTL"))


```

Look at the summarys based on significance:  

```{r}
totalAPA_peakdist_sig=totalAPA_peakdist %>% filter(sig==1)
totalAPA_peakdist_notsig=totalAPA_peakdist %>% filter(sig==0)


summary(totalAPA_peakdist_sig$DisttoPeak)
summary(totalAPA_peakdist_notsig$DisttoPeak)
```

```{r}
ggplot(totalAPA_peakdist, aes(y=DisttoPeak,x=sig, fill=sig, by=sig)) + geom_boxplot()  + scale_fill_discrete(guide = guide_legend(title = "Significant QTL"))
```

Look like there are some outliers that are really far. I will remove variants greater than 1*10^6th away   

```{r}
totalAPA_peakdist_filt=totalAPA_peakdist %>% filter(abs(DisttoPeak) <= 1*(10^6))

ggplot(totalAPA_peakdist_filt, aes(y=DisttoPeak,x=sig, fill=sig, by=sig)) + geom_boxplot()  + scale_fill_discrete(guide = guide_legend(title = "Significant QTL")) + facet_grid(.~strand)

ggplot(totalAPA_peakdist_filt, aes(x=DisttoPeak, fill=sig, by=sig)) + geom_density()  + scale_fill_discrete(guide = guide_legend(title = "Significant QTL")) + facet_grid(.~strand)

```
This gives a similar distribution. 


I did snp - peak. This means if the peak is downstream of the snp on the positive strand the number will be negative. 

In this case the peak is downstream of the snp. 

```{r}
totalAPA_peakdist %>% filter(sig==1) %>% filter(strand=="+") %>%  filter(DisttoPeak < 0) %>% nrow()
totalAPA_peakdist %>% filter(sig==1) %>% filter(strand=="-") %>%  filter(DisttoPeak > 0) %>% nrow()
```
Peak is upstream of the snp. 

```{r}
totalAPA_peakdist %>% filter(sig==1) %>% filter(strand=="+") %>%  filter(DisttoPeak > 0) %>% nrow()
totalAPA_peakdist %>% filter(sig==1) %>% filter(strand=="-") %>%  filter(DisttoPeak < 0) %>% nrow()
```
This means there is about 50/50 distribution around the peak start.  

I am going to plot a violin plot for just the significant ones.  

```{r}
ggplot(totalAPA_peakdist_sig, aes(x=DisttoPeak)) + geom_density()
```
Within 1000 bases of the peak center.  

```{r}
totalAPA_peakdist_sig %>% filter(abs(DisttoPeak) < 1000) %>% nrow()

totalAPA_peakdist_sig %>% filter(abs(DisttoPeak) < 10000) %>% nrow()

totalAPA_peakdist_sig %>% filter(abs(DisttoPeak) < 100000) %>% nrow()
```
29 QTLs are within 1000 bp of the peak center, 57 within 10,000bp and 98 within 100,000bp   



##Expression metrics  


Next I want to pull in the expression values and compare the expression of genes with a total APA qtl in comparison to genes without one.  I will also need to pull in the gene names file to add in the gene names from the ensg ID.   

Remove the # from the file. 

```{r}
expression=read.table("../data/mol_pheno/fastqtl_qqnorm_RNAseq_phase2.fixed.noChr.txt", header = T,stringsAsFactors = F)
expression_mean=apply(expression[,5:73],1,mean,na.rm=TRUE)
expression_var=apply(expression[,5:73],1,var,na.rm=TRUE)
expression$exp.mean= expression_mean 
expression$exp.var=expression_var
expression= expression %>% separate(ID, into=c("Gene.stable.ID", "ver"), sep ="[.]")
```


Now I can pull in the names and join the dataframes.  

```{r}
geneNames=read.table("../data/ensemble_to_genename.txt", sep="\t", header=T,stringsAsFactors = F) 



expression=expression %>% inner_join(geneNames,by="Gene.stable.ID") 

expression=expression %>% select(Chr, start, end, Gene.name, exp.mean,exp.var) %>%  rename("gene"=Gene.name)
```


Now I can join this with the qtls.  

```{r}
totalAPA_wExp=totalAPA %>% inner_join(expression, by="gene") 
```

```{r}
ggplot(totalAPA_wExp, aes(x=exp.mean, by=sig, fill=sig)) + geom_density(alpha=.3)
```

This is not exactly what I want because there are multiple peaks in a gene so some genes are plotted multiple times.  I want to group the QTLs by gene and see if there is 1 sig QTL for that gene.  

```{r}
gene_wQTL= totalAPA_wExp %>% group_by(gene) %>% summarise(sig_gene=sum(as.numeric(as.character(sig)))) %>% ungroup() %>% inner_join(expression, by="gene") %>% mutate(sigGeneFactor=ifelse(sig_gene>=1, 1,0))

gene_wQTL$sigGeneFactor= as.factor(as.character(gene_wQTL$sigGeneFactor))

```
Therea are 92 genes in this set with a QTL.   


```{r}
ggplot(gene_wQTL, aes(x=exp.mean, by=sigGeneFactor, fill=sigGeneFactor)) + geom_density(alpha=.3) +labs(title="Mean in RNA expression by genes with significant QTL", x="Mean in normalized expression") + scale_fill_discrete(guide = guide_legend(title = "Significant QTL"))
```


I can do a similar analysis but test the variance in the gene expression.  

```{r}
ggplot(gene_wQTL, aes(x=exp.var, by=sigGeneFactor, fill=sigGeneFactor)) + geom_density(alpha=.3) + labs(title="Varriance in RNA expression by genes with significant QTL", x="Variance in normalized expression") + scale_fill_discrete(guide = guide_legend(title = "Significant QTL"))
```

### Peak coverage for QTLs 
I can also look at peak coverage for peaks with QLTs and those without. I will first look at this on peak level then mvoe to gene level. The peak scores come from the coverage in the peaks. 

The totalAPA_peakdist data frame has the information I need for this.  

```{r}
ggplot(totalAPA_peakdist, aes(x=PeakScore,fill=sig,by=sig)) + geom_density(alpha=.5)+ scale_x_log10() + labs(title="Peak score by significance") 
```
This is expected. It makes sense that we have more power to detect qtls in higher expressed peaks. This leads me to believe that filtering out low peaks may add power but will not mitigate the effect. 

##Where are the snps:  

Download the GM12878 chromHMM annotation. I downleaded this from uscs and put it in:  

* /Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/GM12878.chromHMM.txt    
* /project2/gilad/briana/genome_anotation_data/GM12878.chromHMM.txt  

Column:  

* bin  
* chrom  
* chromstart  
* chromend  
* name  
* score  
* strand  
* thich start  
* thick end  
* item rgb  


I can make this a bedfile to use a bedtools pipeline:  

* chrom (nochr)  

* start  

* end  

* name (txn hetero ect)   

* score  

* strand  


```{bash,eval=F}
fout = open("/project2/gilad/briana/genome_anotation_data/GM12878.chromHMM.bed",'w')
for ln in open("/project2/gilad/briana/genome_anotation_data/GM12878.chromHMM.txt", "r"):
    bin, chrom, start, end, name, score, strand, thSt, thE, rgb = ln.split()
    chrom=chrom[3:]
    fout.write("%s\t%s\t%s\t%s\t%s\t%s\n"%(chrom, start, end, name, score, strand))
fout.close()
```


```{bash,eval=F}
fout = open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/GM12878.chromHMM.bed",'w')
for ln in open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/GM12878.chromHMM.txt", "r"):
    bin, chrom, start, end, name, score, strand, thSt, thE, rgb = ln.split()
    chrom=chrom[3:]
    fout.write("%s\t%s\t%s\t%s\t%s\t%s\n"%(chrom, start, end, name, score, strand))
fout.close()
```


I also need to create a significant QTL snp bedfile for the total qtls. Bed files are 0 bases meaning I want the end to be the position I care about.  

chrom, start (pos -1), end (pos), name, score (bh), strand

I can do this in python using the /project2/gilad/briana/threeprimeseq/data/perm_APAqtl_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Total_transcript_permResBH.txt. I will make the script general to use on the total or nuclear file.  

QTLres2SigSNPbed.py
```{bash,eval=F}

def main(inFile, outFile):
    fout=open(outFile, "w")
    fin=open(inFile, "r")
    for num, ln in enumerate(fin):
      if num >= 1:
          pid, nvar, shape1, shape2, dummy, sid, dist, npval, slope, ppval, bpval, bh = ln.split()
          chrom, pos= sid.split(":")
          name=sid
          start= int(pos)-1
          end=int(pos)
          strand=pid.split(":")[3].split("_")[1]
          bh=float(bh)
          if bh <= .1: 
              fout.write("%s\t%s\t%s\t%s\t%s\t%s\n"%(chrom, start, end, name, bh, strand))
    fout.close()

if __name__ == "__main__":
    import sys
    fraction=sys.argv[1]
    inFile = "/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_%s_transcript_permResBH.txt"%(fraction)
    outFile= "/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_trans/sigSnps/ApaQTLsignificantSnps_10percFDR_%s.bed"%(fraction)
    main(inFile,outFile) 
```




I am gonig to try to use pybedtools instead for bedtools for this analysis. First I can add it to my conda environment.  


```{bash,eval=F}
import pybedtools 
QTL_SNPs = pybedtools.BedTool('mydata/snps.bed')
```


