---
title: "Use cleanUpdTSeq for internal priming"
author: "Briana Mittleman"
date: "7/24/2018"
output: html_document
---


Install new packages:
```{r, eval=F}
source("https://bioconductor.org/biocLite.R")
biocLite("BSgenome.Hsapiens.UCSC.hg19")
```


Load Packages:  
```{r}
library(workflowr)
library(cleanUpdTSeq)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg19)
library(ggseqlogo)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(tidyr)
```

##Clean Peaks

I am also going to install cleanUpdTSeq on my midway account because I will want to write scripts using this package that can take in any bedfile and will write out the file with the classification results. I can also have the cutoff option be a parameter that will change.  

The test set should have chr, start, end, name, score, strand.  



```{r, eval=F}
#!/bin/rscripts

# usage: ./cleanupdtseq.R in_bedfile, outfile, cuttoff

#this script takes a putative peak file, and output file name and a cuttoff for classification and outputs the file with all of the seqs classified. 

#use optparse for management of input arguments I want to be able to imput the 6up nuc file and write out a filter file  

#script needs to run outside of conda env. should module load R in bash script when I submit it 
library(optparse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cleanUpdTSeq)
library(GenomicRanges)
library(BSgenome.Hsapiens.UCSC.hg19)


option_list = list(
  make_option(c("-f", "--file"), action="store", default=NA, type='character',
              help="input file"),
  make_option(c("-o", "--output"), action="store", default=NA, type='character',
              help="output file"),
  make_option(c("-c", "--cutoff"), action="store", default=NA, type='double',
              help="assignment cuttoff")
)
  

opt_parser <- OptionParser(option_list=option_list)
opt <- parse_args(opt_parser)


#interrupt execution if no file is  supplied
if (is.null(opt$file)){
  print_help(opt_parser)
  stop("Need input file", call.=FALSE)
}

#imput file for test data 
testSet <- read.table(file = opt$file, sep="\t", header=TRUE)
peaks <- BED2GRangesSeq(testSet, withSeq=FALSE)

#build vector with human genome  

testSet.NaiveBayes <- buildFeatureVector(peaks, BSgenomeName=Hsapiens,
                                         upstream=40, downstream=30, 
                                         wordSize=6, alphabet=c("ACGT"),
                                         sampleType="unknown", 
                                         replaceNAdistance=30, 
                                         method="NaiveBayes",
                                         ZeroBasedIndex=1, fetchSeq=TRUE)

#classfy sites with built in classsifer

data(classifier)
testResults <- predictTestSet(testSet.NaiveBayes=testSet.NaiveBayes,
                              classifier=classifier,
                              outputFile=NULL, 
                              assignmentCutoff=opt$cutoff)


#write results  

write.table(testResults, file=opt$output, quote = F, row.names = F, col.names = T)  
```


I will need to module load R in the bash script that writes this.  

```{bash, eval=F}
#!/bin/bash

#SBATCH --job-name=clean_filteredpeakstotal
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=clean_filteredpeakstotal.out
#SBATCH --error=clean_filteredpeakstotal.err
#SBATCH --partition=broadwl
#SBATCH --mem=40G
#SBATCH --mail-type=END


module load R

Rscript cleanupdtseq.R  -f /project2/gilad/briana/threeprimeseq/data/clean.peaks/APAfiltered_named.bed -o /project2/gilad/briana/threeprimeseq/data/clean.peaks/clean_APAfilteredTotal.txt -c .5


```


```{bash, eval=F}
#add names to bed file with peaks 
#awk '{print $1 "\t" $2 "\t" $3 "\t" $1 ":" $2 ":" $3 "\t"  $4 "\t"  $5 "\t" $6}' APAfiltered.bed > APAfiltered_named.bed


seq 1 199880 > peak.num.txt
paste APAfiltered.bed peak.num.txt | column -s $'\t' -t > temp
awk '{print $1 "\t" $2 "\t" $3 "\t" $7  "\t"  $4 "\t"  $5 "\t" $6}' temp >  APAfiltered_named.bed

```


##Characterize results  

This cuttoff results in a move from 199880 to 125825 called sites.  


```{r}
peaks=read.table("../data/clean_peaks/clean_APAfilteredTotal.txt", header = T, stringsAsFactors = F)
```

Plot the density of the probabilities. I expect a bimodal distribution.  

```{r}
ggplot(peaks, aes(probTrue)) + geom_density(fill="blue") + labs(title="Density of Probability Peak is a True APA peak", x="Probability True PAS")
```

Next I want to make logo plots for the upstream sequences seperated by class. I expect to see an overrepresentation of A/T in the upstream of the false samples.  

```{r}
true_peaks=peaks %>% filter(pred.class==1)
false_peaks=peaks %>% filter(pred.class==0)
```

I can extract just the upstream seq for each class.  

```{r}
true_peaks_up=peaks %>% filter(pred.class==1) %>% select(UpstreamSeq)
false_peaks_up= peaks %>% filter(pred.class==0) %>% select(UpstreamSeq)


```

###Sequence composition  
```{r}
trueplot_up=ggseqlogo(true_peaks_up,seq_type='dna', method = 'prob') + labs(x="Base number", title="Upstream Seq: True PAS")

falseplot_up=ggseqlogo(false_peaks_up,seq_type='dna', method = 'prob')  + labs(x="Base number", title="Upstream Seq: False PAS")


gridExtra::grid.arrange(trueplot_up,falseplot_up)
```

I can do the same thing for the downstream seq.  

```{r}
true_peaks_down=peaks %>% filter(pred.class==1) %>% select(DownstreamSeq)
false_peaks_down= peaks %>% filter(pred.class==0) %>% select(DownstreamSeq)


trueplot_down=ggseqlogo(true_peaks_down,seq_type='dna', method = 'prob') + labs(x="Base number", title="Downstream Seq: True PAS")

falseplot_down=ggseqlogo(false_peaks_down,seq_type='dna', method = 'prob')  + labs(x="Base number", title="Downstream Seq: False PAS")


gridExtra::grid.arrange(trueplot_down,falseplot_down)
```

###Length differnces between true and false peaks

I can join all of the information from the original bed with the results using a join by the name.

```{r}
names=c("chr", "start", "end", "PeakName", "Cov", "Strand", "score")
YL_peaks=read.table("../data/clean_peaks/APAfiltered_named.bed", col.names = names)
```

```{r}
full_peaks= inner_join(YL_peaks, peaks, by="PeakName") %>% mutate(length=end-start)
full_peaks$pred.class= as.factor(full_peaks$pred.class)
```


```{r}

ggplot(full_peaks, aes(length, group=pred.class, fill=pred.class)) + geom_density(alpha=.4) + scale_x_log10() + labs(title="Peak lengths do not differ by predicted class", x="Length of called Peak") + scale_fill_manual(values=c("red", "blue"), name="Predicted Class", labels=c("False Positive", "True PAS"))
```



###Coverage differnces between true and false peaks
```{r}

ggplot(full_peaks, aes(x=Cov, group=pred.class, fill=pred.class)) + geom_density(alpha=.4) + scale_x_log10() + labs(title="Peak coverage by predicted class", x=" Peak coverage") + scale_fill_manual(values=c("red", "blue"), name="Predicted Class", labels=c("False Positive", "True PAS"))
```
```


