---
title: "Condition on Types ApaQTLs"
author: "Briana Mittleman"
date: "10/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I want to use the resuls from the total and nuclear APA qtl calling. I will ask if conditioning on a nuclear QTL increases the signal in the total QTL and vice versa. I will start with the significant snp-peak pairs from the permuted files. I will then overlap with the nominal pvalues from the other fraction. I will do this similar to how I did in the overlaMolQTL analysis. However in this analysis I do not have the multiple peaks per gene problem that I have when I overlap the pvalues. I can map the same peak to snp pair.  

Due to file size I will do this only with the permuted files.  


* Load Libraries  

```{r}
library(tidyverse)
library(reshape2)
library(workflowr)
```

* Load Data 

**Permuted**
```{r}
permTot=read.table("../data/perm_QTL_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Total_transcript_permResBH.txt",header=T ,stringsAsFactors = F)
permNuc=read.table("../data/perm_QTL_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Nuclear_transcript_permResBH.txt",header=T, stringsAsFactors = F)
```


**Nominal**  

```{r}
nomnames=c("pid", "sid", "dist", "npval", "slope")
#nomTot=read_table("../data/nom_QTL_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Total_NomRes.txt", col_names=nomnames, col_types = c(col_character(), col_character(), col_double(), col_double(), col_double()))
#nomNuc=read_table("../data/nom_QTL_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Nuclear_NomRes.txt", col_names=nomnames, col_types = c(col_character(), col_character(), col_double(), col_double(), col_double()))
```



##Make QQplots  

```{r}
overlapQTLplot_totalQTL=function(cut, plotfile){

    #helper functions  
    sigsnp=function(cutoff){
      permTot$bh=p.adjust(permTot$bpval, method="fdr")
      file_sig=permTot %>% filter(-log10(bh)> cutoff) %>% select(sid)
      print(paste("Sig snps=", nrow(file_sig), sep=" "))
      return(file_sig)
    }
    randomsnps=function(SigSnpList){
      nsnp=nrow(SigSnpList)
      randomSnpDF= permTot %>% sample_n(nsnp) %>% arrange(sid) %>% select(sid)
      return(randomSnpDF)
    }
    top_Nuclear=function(snp_list){
      filt_nuc=permNuc %>% semi_join(snp_list, by="sid") %>% group_by(sid) %>% add_tally() %>% ungroup() %>% mutate(corrPval=bpval)
      filt_nuc_top= filt_nuc %>% group_by(sid) %>% top_n(-1, corrPval)
      print(paste("Nuclear overlap=", nrow(filt_nuc_top), sep=" "))
      return(filt_nuc_top)
    }
    makeQQ=function(test, baseline){
      plot=qqplot(-log10(runif(nrow(baseline))), -log10(baseline$corrPval), ylab="Observed", xlab="Expected", main="Significant Total QTLs- nuclear Pval")
      points(sort(-log10(runif(nrow(test)))), sort(-log10(test$corrPval)), col= alpha("Red"))
      abline(0,1)
      return(plot)
    }
  
    TL=sigsnp(cut)
    BL=randomsnps(TL)
    #top snps test and base total
    topN_T=top_Nuclear(TL)
    topN_B=top_Nuclear(BL)
    #plot Total
    png(plotfile)
    totalPlot=makeQQ(topN_T,topN_B)
    dev.off()

}

overlapQTLplot_totalQTL(1, "../output/plots/TotalQTLinNuclear.png")

```


```{r}


overlapQTLplot_totalQTL=function(cut, plotfile){

    #helper functions  
    sigsnp=function(cutoff){
      permNuc$bh=p.adjust(permNuc$bpval, method="fdr")
      file_sig=permNuc %>% filter(-log10(bh)> cutoff) %>% select(sid)
      print(paste("Sig snps=", nrow(file_sig), sep=" "))
      return(file_sig)
    }
    randomsnps=function(SigSnpList){
      nsnp=nrow(SigSnpList)
      randomSnpDF= permNuc %>% sample_n(nsnp) %>% arrange(sid) %>% select(sid)
      return(randomSnpDF)
    }
    top_Total=function(snp_list){
      filt_tot=permTot %>% semi_join(snp_list, by="sid") %>% group_by(sid) %>% add_tally() %>% ungroup() %>% mutate(corrPval=bpval)
      filt_tot_top= filt_tot %>% group_by(sid) %>% top_n(-1, corrPval)
      print(paste("Total overlap=", nrow(filt_tot_top), sep=" "))
      return(filt_tot_top)
    }
    makeQQ=function(test, baseline){
      plot=qqplot(-log10(runif(nrow(baseline))), -log10(baseline$corrPval), ylab="Observed", xlab="Expected", main="Significant Nuclear QTLs- Total Pval")
      points(sort(-log10(runif(nrow(test)))), sort(-log10(test$corrPval)), col= alpha("Red"))
      abline(0,1)
      return(plot)
    }
  
    TL=sigsnp(cut)
    BL=randomsnps(TL)
    #top snps test and base total
    topN_T=top_Total(TL)
    topN_B=top_Total(BL)
    #plot Total
    png(plotfile)
    totalPlot=makeQQ(topN_T,topN_B)
    dev.off()

}
overlapQTLplot_totalQTL(1, "../output/plots/NuclearQTLinTotal.png")
```


I should change this to focus on peak. I can say give me the genes with significant QTLs in total or nuclear then look at the pvalues for those peaks in the other file. As I did before, I am going to work on all of the functions seperatly then put them together.  

Get the peaks with significant QTLs, and the same number of random peaks.   


```{r, eval=F}

sigpeak=function(cutoff){
    permNuc$bh=p.adjust(permNuc$bpval, method="fdr")
    file_sig=permNuc %>% filter(-log10(bh)> cutoff) %>%separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
    print(paste("Sig peaks=", nrow(file_sig), sep=" "))
    return(file_sig)
    }
    
x=sigpeak(1)

randompeak=function(SigSnpList){
  nsnp=nrow(SigSnpList)
  randomPeakDF= permNuc %>% sample_n(nsnp) %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
  return(randomPeakDF)
}

y=randompeak(x)

```
I can now get the top pvalue for each of these using the total permuted pval.  

```{r,eval=F}

Peak_overlap=function(snp_list){
      filt_tot=permTot %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% semi_join(snp_list, by="peak")
      print(paste("Total overlap=", nrow(filt_tot), sep=" "))
      return(filt_tot)
}


#run on real sig peaks and random peaks  
Test=Peak_overlap(x)
base=Peak_overlap(y)

```

Plot:  


```{r,eval=F}
  makeQQ_peak=function(test, baseline){
      plot=qqplot(-log10(runif(nrow(test))), -log10(test$bpval), ylab="Observed", xlab="Expected", main="Peaks with Significant QTLs in Nuc \n pvalues in Tot")
      points(sort(-log10(runif(nrow(baseline)))), sort(-log10(baseline$bpval)), col=alpha("Red"))
      abline(0,1)
      return(plot)
}
plot=makeQQ_peak(Test,base)



```


```{r}
SigNucPeakOverlapTot=function(cutoff, plotfile){
    sigpeak=function(cutoff){
      permNuc$bh=p.adjust(permNuc$bpval, method="fdr")
      file_sig=permNuc %>% filter(-log10(bh)> cutoff) %>%separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
      print(paste("Sig peaks=", nrow(file_sig), sep=" "))
      return(file_sig)
      }
  
  randompeak=function(SigSnpList){
    nsnp=nrow(SigSnpList)
    randomPeakDF= permNuc %>% sample_n(nsnp) %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
    return(randomPeakDF)
  }
  Peak_overlap=function(snp_list){
        filt_tot=permTot %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% semi_join(snp_list, by="peak")
        print(paste("Total overlap=", nrow(filt_tot), sep=" "))
        return(filt_tot)
  }

  makeQQ_peak=function(test, baseline){
      plot=qqplot(-log10(runif(nrow(baseline))), -log10(baseline$bpval), ylab="Observed", xlab="Expected", main="Peaks with Significant QTLs in Nuclear \n pvalues in Total")
      points(sort(-log10(runif(nrow(test)))), sort(-log10(test$bpval)), col= alpha("Red"))
      abline(0,1)
      return(plot)
  }
  
  
testPeaks=sigpeak(1)
basePeaks=randompeak(testPeaks)
testSet=Peak_overlap(testPeaks)
baselineSet=Peak_overlap(basePeaks)
png(plotfile)
plot=makeQQ_peak(testSet,baselineSet )
dev.off()
}

SigNucPeakOverlapTot(1, "../output/plots/SigNucPeakTotpval.png")
```



```{r}
SigTotPeakOverlapNuc=function(cutoff, plotfile){
    sigpeak=function(cutoff){
      permTot$bh=p.adjust(permTot$bpval, method="fdr")
      file_sig=permTot %>% filter(-log10(bh)> cutoff) %>%separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
      print(paste("Sig peaks=", nrow(file_sig), sep=" "))
      return(file_sig)
      }
  
  randompeak=function(SigSnpList){
    nsnp=nrow(SigSnpList)
    randomPeakDF= permTot %>% sample_n(nsnp) %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% select(peak)
    return(randomPeakDF)
  }
  Peak_overlap=function(snp_list){
        filt_nuc=permNuc %>% separate(pid, sep = ":", into=c("chr", "start", "end", "id")) %>% separate(id, sep = "_", into=c("gene", "strand", "peak")) %>% semi_join(snp_list, by="peak")
        print(paste("Nuclear overlap=", nrow(filt_nuc), sep=" "))
        return(filt_nuc)
  }

  makeQQ_peak=function(test, baseline){
      plot=qqplot(-log10(runif(nrow(baseline))), -log10(baseline$bpval), ylab="Observed", xlab="Expected", main="Peaks with Significant QTLs in Total \n pvalues in Nuclear")
      points(sort(-log10(runif(nrow(test)))), sort(-log10(test$bpval)), col= alpha("Red"))
      abline(0,1)
      return(plot)
  }
  
  
testPeaks=sigpeak(1)
basePeaks=randompeak(testPeaks)
testSet=Peak_overlap(testPeaks)
baselineSet=Peak_overlap(basePeaks)
png(plotfile)
plot=makeQQ_peak(testSet,baselineSet )
dev.off()
}

SigTotPeakOverlapNuc(1, "../output/plots/SigTotPeakNucpval.png")
```
```

