---
title: "cov.200bp.wind"
author: "Briana Mittleman"
date: "5/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
I will use this analysis to bin the genome into 200bp windows and look at coverage for the 3' seq libraries for each of these windows. I will use this data then in the leafcutter pipeline to look at differences between data from the total and nuclear fractions.  

I performed a similar analysis for the net-seq data so some of the code will come from that.  https://brimittleman.github.io/Net-seq/create_blacklist.html  

###Map reads to bins  
The binned genome file is called: genome_200_wind_fix2.saf, it is in my genome annotation directory.  

```{bash, eval=FALSE}
#!/bin/bash

#SBATCH --job-name=cov200
#SBATCH --time=8:00:00
#SBATCH --output=cov200.out
#SBATCH --error=cov200.err
#SBATCH --partition=broadwl
#SBATCH --mem=20G
#SBATCH --mail-type=END

module load Anaconda3  

source activate three-prime-env

#input is a bam 
sample=$1


describer=$(echo ${sample} | sed -e 's/.*\YL-SP-//' | sed -e "s/-sort.bam$//")



featureCounts -T 5 -a /project2/gilad/briana/genome_anotation_data/an.int.genome_200_strandspec.saf -F 'SAF' -o /project2/gilad/briana/threeprimeseq/data/cov_200/${describer}_FC200.cov.bed $1
```


I will need to create a wrapper to run this for all of the files. 


```{bash, eval=FALSE}
#!/bin/bash

#SBATCH --job-name=w_cov200
#SBATCH --time=8:00:00
#SBATCH --output=w_cov200.out
#SBATCH --error=w_cov200.err
#SBATCH --partition=broadwl
#SBATCH --mem=8G
#SBATCH --mail-type=END


for i in $(ls /project2/gilad/briana/threeprimeseq/data/sort/*.bam); do
            sbatch cov200.sh $i 
        done
```

Current analysis is not stand specific. I need to make windows for the negative strand. To do this I need to copy the genome_200_wind_fix2.saf file but with geneIDs starting with the last number of the file and with a - for the strand. The last window number is 15685849. I will have to start from 15685850. 

In general I will use awk to create the file.  The last number is 31371698 because that is 2 * the number of bins in the genome. I w 

```{bash, eval=F}
#i will delete the top line at the end
seq 15685849 31371698 > neg.bin.num.txt

 cut -f1 neg.bin.num.txt | paste - genome_200_wind_fix2.saf | awk  '{ if (NR>1) print $1 "\t" $3 "\t" $4 "\t" $5 "\t" "-"}' >  genome_200_wind_fix2.negstrand.saf 

#cat files together  

cat genome_200_wind_fix2.saf  genome_200_wind_fix2.negstrand.saf  > genome_200_strandsspec_wind.saf
 
```


I can use this to get coverage in all of the windows with strand specificity.  I will call this script ss_cov200.sh

```{bash, eval=FALSE}
#!/bin/bash

#SBATCH --job-name=sscov200
#SBATCH --time=8:00:00
#SBATCH --output=sscov200.out
#SBATCH --error=sscov200.err
#SBATCH --partition=broadwl
#SBATCH --mem=20G
#SBATCH --mail-type=END

module load Anaconda3  

source activate three-prime-env

#input is a bam 
sample=$1


describer=$(echo ${sample} | sed -e 's/.*\YL-SP-//' | sed -e "s/-sort.bam$//")



featureCounts -T 5 -s 1 -O --fraction -a /project2/gilad/briana/genome_anotation_data/genome_200_strandsspec_wind.saf -F 'SAF' -o /project2/gilad/briana/threeprimeseq/data/ss_cov200/${describer}_ssFC200.cov.bed $1
```

Try this with. /project2/gilad/briana/threeprimeseq/data/sort/YL-SP-18486-N_S10_R1_001-sort.bam 

I will update my wrapper to use this script.  


The current script does not allow reads that map to multiple bins. We expect then so I will update the featureCounts code to account for this. 

-O allows multi mapping
-fraction will put a fraction of the read in each bin

The next step is to add genes annotations to each bin. I will do this with bedtools closest on my window file.  

gene file: /project2/gilad/briana/genome_anotation_data/gencode.v19.annotation.proteincodinggene.sort.bed

I want to keep the windows with gene and add the name of the gene they are in.  

a= windows
b= genes 

force stranded= -s

I need to make the window file a sorted bed file. It should be the chr number without the 'chr' tag, start, end, bin number, ".", strand.  

```{bash, eval=F}
awk '{if (NR>1) print $2 "\t" $3 "\t" $4 "\t" $1 "\t" "." "\t" $5}' genome_200_strandsspec_wind.saf  | sed 's/^chr//' | sort -k1,1 -k2,2n > genome_200_strandspec.bed

```


```{bash, eval=F}
#!/bin/bash

#SBATCH --job-name=annotate_wind
#SBATCH --time=8:00:00
#SBATCH --output=an_wind.out
#SBATCH --error=an_wind.err
#SBATCH --partition=broadwl
#SBATCH --mem=30G
#SBATCH --mail-type=END

module load Anaconda3

source activate three-prime-env


bedtools closest -s -a genome_200_strandspec.bed -b gencode.v19.annotation.proteincodinggene.sort.bed > annotated.genome_200_strandspec.bed
```

Now i can use intersect to only keep the windows that interdect that protien coding genes.  
```{bash, eval=F}
#!/bin/bash

#SBATCH --job-name=int_wind
#SBATCH --time=8:00:00
#SBATCH --output=int_wind.out
#SBATCH --error=int_wind.err
#SBATCH --partition=broadwl
#SBATCH --mem=30G
#SBATCH --mail-type=END

module load Anaconda3

source activate three-prime-env


bedtools intersect -wa -sorted -s -a annotated.genome_200_strandspec.bed -b gencode.v19.annotation.proteincodinggene.sort.bed > annotated.int.genome_200_strandspec.bed
```


```{bash, eval=F}
awk '{print $1 "\t" $2 "\t" $3 "\t" $4 "\t" $5 "\t" $6 "\t" $10}'  annotated.int.genome_200_strandspec.bed >  an.int.genome_200_strandspec.bed
```

I went from 31590487 to 7371747 windows. I need to make this into a saf file and the name of the window will be the number.gene

```{bash, eval=F}
awk '{print $4"."$7 "\t" $1 "\t" $2 "\t" $3 "\t" $6}'  an.int.genome_200_strandspec.bed > an.int.genome_200_strandspec.saf

#go into the file with vi and add header
```
Now I can change my feature counts script to use this file instead.  


I need to get rid of the lines with 2 genes overlapping in the bin. I will do this by removing the lines with a :.  
```{bash, eval=F}

for i in $(ls *.bed); do
      cat $i | grep -v -e ";" > ../ss_cov200_no_overlap/$i
  done

```

The  next step is to bind all of these files.  This file will have all 6323877 windows as the rows and columns for each of the 32 files  

```{bash, eval=F}

less 18486-N_S10_R1_001_ssFC200.cov.bed  | cut -f1-6 > tmp 
for i in ./*cov.bed; do
echo "$i"
less ${i} | cut -f7 >col
paste tmp col> tmp2; mv tmp2 tmp; rm col; done

mv tmp ssFC200.cov.bed
```

This in now ready to move to R an work with it here.  


###Assess bin coverage

```{r}
library(workflowr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(edgeR)
library(reshape2)
```




```{r}
names=c("N_18486","T_18486","N_18497","T_18497","N_18500","T_18500","N_18505",'T_18505',"N_18508","T_18508","N_18853","T_18853","N_18870","T_18870","N_19128","T_19128","N_19141","T_19141","N_19193","T_19193","N_19209","T_19209","N_19223","N_19225","T_19225","T_19223","N_19238","T_19238","N_19239","T_19239","N_19257","T_19257")
```

```{r}
cov_all=read.table("../data/ssFC200.cov.bed", header = T, stringsAsFactors = FALSE)
#remember name switch!
names=c("Geneid","Chr", "Start", "End", "Strand", "Length", "N_18486","T_18486","N_18497","T_18497","N_18500","T_18500","N_18505",'T_18505',"N_18508","T_18508","N_18853","T_18853","N_18870","T_18870","N_19128","T_19128","N_19141","T_19141","N_19193","T_19193","N_19209","T_19209","N_19223","N_19225","T_19225","T_19223","N_19238","T_19238","N_19239","T_19239","N_19257","T_19257")

colnames(cov_all)= names
```

Plot the density of the log of the counts.  

```{r}
cov_nums_only=cov_all[,7:38]
cov_nums_only_log=log10(cov_nums_only)
plotDensities(cov_nums_only_log,legend = "bottomright", main="bin log 10 counts")
```
Now I want to filter for bins that have 0 reads in >16 samples.  


```{r}
keep.exprs=rowSums(cov_nums_only>0) >= 16

cov_all_filt=cov_all[keep.exprs,]
bin.genes=cov_all_filt[,1]
```

I will now look at the densities.

```{r}
cov_all_filt_log=log10(cov_all_filt[,7:38] + 1) 

plotDensities(cov_all_filt_log,legend = "bottomright", main="Filtered bin log10 +1 counts")
```

I want to make boxplots for each of these lines. I should tidy the data with a column for total or nuclear.  

```{r}
sample=c("N_18486","T_18486","N_18497","T_18497","N_18500","T_18500","N_18505",'T_18505',"N_18508","T_18508","N_18853","T_18853","N_18870","T_18870","N_19128","T_19128","N_19141","T_19141","N_19193","T_19193","N_19209","T_19209","N_19223","N_19225","T_19225","T_19223","N_19238","T_19238","N_19239","T_19239","N_19257","T_19257")
fraction=c("N","T","N","T","N","T","N",'T',"N","T","N","T","N","T","N","T","N","T","N","T","N","T","N","N","T","T","N","T","N","T","N","T")

cov_all_filt_log_gen=cbind(bin.genes,cov_all_filt_log)

cov_all_tidy= cov_all_filt_log_gen%>% gather(sample, value, -bin.genes)


#add fraction column

cov_all_tidy_frac=cov_all_tidy %>% mutate(fraction=ifelse(grepl("T",sample), "total", "nuclear")) %>% mutate(line=substr(sample,3,7))

```
Make a heatmap: 

```{r}
bin_count=ggplot(cov_all_tidy_frac, aes(x = line, y=value,fill=fraction )) + geom_boxplot(position="dodge") + labs(y="log10 count + 1", title="Bins in nuclear fractions have larger counts " )
bin_count

#ggsave("../output/plots/bin_counts_by_line.png", bin_count)
```


###Non-zero bins  

For the next section of this analysis I want to look at how many of the bins have non zero counts. I will do this over all then I will gather per gene and look at this. I will use the filtered non transformed data.  

```{r}
cov_all_filt_genes=separate(data = cov_all_filt, col = Geneid, into = c("bin", "gene"), sep = ".E") 
cov_all_filt_genes$gene= paste( "E",  cov_all_filt_genes$gene, sep="" )
cov_all_filt_num=cov_all_filt_genes[,8:39]
non_zero=colSums(cov_all_filt_num != 0)

#make a data frame to plot this

non_zero_df=data.frame(non_zero)
non_zero_df= non_zero_df %>% mutate(fraction=ifelse(grepl("T",rownames(non_zero_df)), "total", "nuclear")) %>% mutate(line=substr(rownames(non_zero_df),3,7))


non_zero_plot=ggplot(non_zero_df, aes(x = line, y=non_zero, fill=fraction )) + geom_bar(position="dodge",stat="identity") + labs(y="Non zero bins", title="Number of bins with reads after filtering")
non_zero_plot
#ggsave("../output/plots/non_zero_bins.png", non_zero_plot)
```

This analysis is bins over all. I want to look at this by gene. I want to get a number of nonzero bins per gene/ number of bins for that gene. I will use the gather function.  

```{r}
cov_all_filt_small=cbind(cov_all_filt_genes[,1:2],cov_all_filt_genes[,8:39])
cov_all_filt_pergene=cov_all_filt_small %>% gather(sample, value, -gene, -bin)  %>% group_by(gene, sample) %>% summarise(non_zero=sum(value!=0)/n())%>% mutate(fraction=ifelse(grepl("T",sample), "total", "nuclear")) %>% mutate(line=substr(sample,3,7))

```

Now I have the number of non zero bins in that gene/ number of bins in that gene. I need to think about the way to plot this.  


###Prepare data for leafcutter: 

For leafcutter I need the data to be reads in one bin compared to other bins in that gene.
