---
title: "eRNAs and Repetitive elements"
author: "Briana Mittleman"
date: "3/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(workflowr)
library(tidyverse)
library(cowplot)
library(reshape2)
```

##Repetitive elements:  
Could this be do to repetitive elements  

Process: 
/project2/gilad/briana/genome_anotation_data/RepeatMask.dms  

I just need to cut the chr to make the chroms the same as mine  


```{bash,eval=F}
sed 's/^chr//'  /project2/gilad/briana/genome_anotation_data/RepeatMask.dms   > /project2/gilad/briana/genome_anotation_data/RepeatMask.bed

```



BothFracDTPlotRepeats_noMPFilt.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=BothFracDTPlotRepeats_noMPFilt
#SBATCH --account=pi-yangili1
#SBATCH --time=36:00:00
#SBATCH --output=BothFracDTPlotRepeats_noMPFilt.out
#SBATCH --error=BothFracDTPlotRepeats_noMPFilt.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


computeMatrix reference-point -S /project2/gilad/briana/threeprimeseq/data/mergedBW/Total_MergedBamCoverage.bw /project2/gilad/briana/threeprimeseq/data/mergedBW/Nuclear_MergedBamCoverage.bw -R /project2/gilad/briana/genome_anotation_data/RepeatMask.bed -b 1000 -a 1000  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_Repetitive_Nompfilt.gz

plotHeatmap --sortRegions descend -m /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_Repetitive_Nompfilt.gz --refPointLabel "Repetitive Regions" --plotTitle "Combined Reads at Repetitive Regions" --heatmapHeight 7 --colorMap YlGnBu  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_Repetitive_Nompfilt.png
```


##eRNA 

download LCL eRNAs from phantom  

http://enhancer.binf.ku.dk/presets/  

process this file 

/project2/gilad/briana/genome_anotation_data/0000945_lymphocyte_of_B_lineage_differentially_expressed_enhancers.bed


interactively in python
```{bash,eval=F}
inFile=open("/project2/gilad/briana/genome_anotation_data/CL:0000945_lymphocyte_of_B_lineage_differentially_expressed_enhancers.bed", "r")
outBed=open("/project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed","w")

for ln in inFile:
    chrom=ln.split()[0]
    chromnoch=chrom[3:]
    start=int(ln.split()[1])
    end=int(ln.split()[2])  
    outBed.write("%s\t%d\t%d\n"%(chromnoch, start,end))
outBed.close()
```



Look at this in total and nuclear three prime seq BW  

BothFracDTPloteRNA_noMPFilt.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=BothFracDTPloteRNA_noMPFilt
#SBATCH --account=pi-yangili1
#SBATCH --time=36:00:00
#SBATCH --output=BothFracDTPloteRNA_noMPFilt.out
#SBATCH --error=BothFracDTPloteRNA_noMPFilt.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


computeMatrix reference-point -S /project2/gilad/briana/threeprimeseq/data/mergedBW/Total_MergedBamCoverage.bw /project2/gilad/briana/threeprimeseq/data/mergedBW/Nuclear_MergedBamCoverage.bw -R /project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed -b 1000 -a 1000  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt.gz

plotHeatmap --sortRegions descend -m /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt.gz --refPointLabel "eRNA Regions" --plotTitle "Combined Reads at eRNA" --heatmapHeight 7 --colorMap YlGnBu  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt.png
```



Do this as region rather than reference point  
BothFracDTPloteRNA_noMPFilt_region.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=BothFracDTPloteRNA_noMPFilt_region
#SBATCH --account=pi-yangili1
#SBATCH --time=36:00:00
#SBATCH --output=BothFracDTPloteRNA_noMPFilt_region.out
#SBATCH --error=BothFracDTPloteRNA_noMPFilt_region.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


computeMatrix scale-regions -S /project2/gilad/briana/threeprimeseq/data/mergedBW/Total_MergedBamCoverage.bw /project2/gilad/briana/threeprimeseq/data/mergedBW/Nuclear_MergedBamCoverage.bw -R /project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed -b 500 -a 500  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt_regions.gz

plotHeatmap --sortRegions descend -m /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt_regions.gz --refPointLabel "eRNA Regions" --plotTitle "Combined Reads at eRNA" --heatmapHeight 7 --colorMap YlGnBu  -out /project2/gilad/briana/threeprimeseq/data/LianoglouDeepTools/BothFrac_eRNA_Nompfilt_regions.png
```

###overlap potential drivers of extra peaks  
Does not look like there are a strong driver. I will see if any of these overlap with our peaks.I will need to look at the opposite strand overlap or use the fixed strand peaks. I will ask how many of these eRNAs or rep elements overlap a peak.  


I want to run the overlap in all of the peaks as well as those that have been filtered 5%  


/project2/gilad/briana/threeprimeseq/data/mergedPeaks_noMP_GeneLoc/Filtered_APApeaks_merged_allchrom_noMP.sort.named.noCHR_geneLoc.bed

/project2/gilad/briana/threeprimeseq/data/peaks4DT/Peaks_5percCov_fixedStrand.bed  

fix strand for nonfiltered:  

fixStrand4DTplots_allpeaks.py

```{bash,eval=F}
peaksIn="/project2/gilad/briana/threeprimeseq/data/mergedPeaks_noMP_GeneLoc/Filtered_APApeaks_merged_allchrom_noMP.sort.named.noCHR_geneLoc.bed"
PeakOut="/project2/gilad/briana/threeprimeseq/data/peaks4DT/APAPeaks_allPeaks_fixedStrand.bed"


def fix_strand(Fin,Fout):
    fout=open(Fout,"w")
    for ln in open(Fin, "r"):
        chrom, start, end, name, score, strand, score2, pos = ln.split()
        if strand=="+":
            nameF="peak" + name + ":" + pos
            fout.write("%s\t%s\t%s\t%s\t%s\t-\n"%(chrom,start,end,nameF,score))
        else:
            nameF="peak" + name + ":" + pos
            fout.write("%s\t%s\t%s\t%s\t%s\t+\n"%(chrom,start,end,nameF,score))
    fout.close()
    
    
fix_strand(peaksIn, PeakOut)

```

ernas: /project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed 
repetitive elements: /project2/gilad/briana/genome_anotation_data/RepeatMask.bed

make a python script with pybedtools that will take any bed file and overlap it 

overlapWFilteredPeaks.py
```{bash,eval=F}
def main(infile, outfile):
    peak_file=open("/project2/gilad/briana/threeprimeseq/data/peaks4DT/APAPeaks_5percCov_fixedStrand.bed","r")
    peak=pybedtools.BedTool(peak_file)
    elementFile=open(infile, "r")
    for i,ln in enumerate(elementFile):
       if i == 0:
           if len(ln.split()) > 3:
               strand= "yes"
           else:
               strand= "no"
       else:
          break
    print(strand)
    elements=pybedtools.BedTool(elementFile)
    if strand== "yes": 
        elemOverpeak=elements.intersect(peak, wa=True,wb=True, s=True)
    else:
        elemOverpeak=elements.intersect(peak, wa=True,wb=True)
    elemOverpeak.saveas(outfile)

if __name__ == "__main__":
    import sys
    import pybedtools
    infile=sys.argv[1]
    outfile=sys.argv[2]
    main(infile, outfile)
```

run:  

```{bash,eval=F}
python overlapWFilteredPeaks.py /project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed  /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/FilteredPeak_overeRNA.txt  

python overlapWFilteredPeaks.py /project2/gilad/briana/genome_anotation_data/RepeatMask.bed /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/FilteredPeak_overRepElements.txt  
```


overlapWAllPeaks.py
```{bash,eval=F}
def main(infile, outfile):
    peak_file=open("/project2/gilad/briana/threeprimeseq/data/peaks4DT/APAPeaks_allPeaks_fixedStrand.bed","r")
    peak=pybedtools.BedTool(peak_file)
    elementFile=open(infile, "r")
    for i,ln in enumerate(elementFile):
       if i == 0:
           if len(ln.split()) > 3:
               strand= "yes"
           else:
               strand= "no"
       else:
          break
    print(strand)
    elements=pybedtools.BedTool(elementFile)
    if strand== "yes": 
        elemOverpeak=elements.intersect(peak, wa=True,wb=True, s=True)
    else:
        elemOverpeak=elements.intersect(peak, wa=True,wb=True)
    elemOverpeak.saveas(outfile)

if __name__ == "__main__":
    import sys
    import pybedtools
    infile=sys.argv[1]
    outfile=sys.argv[2]
    main(infile, outfile)
    
```

run:  

```{bash,eval=F}
python overlapWAllPeaks.py /project2/gilad/briana/genome_anotation_data/LCLenhancerRNA.bed  /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/AllPeak_overeRNA.txt  

python overlapWAllPeaks.py /project2/gilad/briana/genome_anotation_data/RepeatMask.bed /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/AllPeak_overRepElements.txt  
```


How long are each of these

Full eRNA file:  1167
All peak eRNA: 128 
Filt peak eRNA: 14  

Full rep file: 5298130
All peak rep:52965
Filt peak rep: 9542  

 
For Repetitive seq. I can pull this in here and look at the proportion in each type.  

```{r}
repEl=read.table("../data/FeatureoverlapPeaks/FilteredPeak_overRepElements.txt", col.names = c("repCHR", "repStart", "repEnd", "Type", "length", "repStrand", "peakChr", "peakStart", "peakEnd", "peak", "score", "peakStrand"),stringsAsFactors = F)

repEl_sum= repEl %>% group_by(Type) %>% summarise(NperType=n()) %>% mutate(Proppertype=NperType/nrow(repEl))
```

```{r}
ggplot(repEl_sum, aes(x=Type,y=Proppertype) ) + geom_bar(stat="identity") +  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
summary(repEl_sum$Proppertype)


```


To get an expectation for this I will shuffle my peaks around the genome and overlap again.  

Do this interactively  
```{bash,eval=F}
import pybedtools 
peaks= pybedtools.BedTool("/project2/gilad/briana/threeprimeseq/data/peaks4DT/APAPeaks_5percCov_fixedStrand.bed")
peaks_shuf=peaks.shuffle(genome='hg19')
peaks_shuf.saveas("/project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/shuffled_FilterPeaks.bed")

```


```{bash,eval=F}
sort -k1,1 -k2,2n /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/shuffled_FilterPeaks.bed | sed 's/^chr//'  > /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/shuffled_FilterPeaks.sort.bed
```


overlapWShuffledPeaks.py
```{bash,eval=F}
def main(infile, outfile):
    peak_file=open("/project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/shuffled_FilterPeaks.sort.bed","r")
    peak=pybedtools.BedTool(peak_file)
    elementFile=open(infile, "r")
    for i,ln in enumerate(elementFile):
       if i == 0:
           if len(ln.split()) > 3:
               strand= "yes"
           else:
               strand= "no"
       else:
          break
    print(strand)
    elements=pybedtools.BedTool(elementFile)
    if strand== "yes": 
        elemOverpeak=elements.intersect(peak, wa=True,wb=True, s=True)
    else:
        elemOverpeak=elements.intersect(peak, wa=True,wb=True)
    elemOverpeak.saveas(outfile)

if __name__ == "__main__":
    import sys
    import pybedtools
    infile=sys.argv[1]
    outfile=sys.argv[2]
    main(infile, outfile)
```

```{bash,eval=F}
python overlapWShuffledPeaks.py /project2/gilad/briana/genome_anotation_data/RepeatMask.bed /project2/gilad/briana/threeprimeseq/data/FeatureoverlapPeaks/ShuffledPeak_overRepElements.txt  
```

```{r}
repEl_shuf=read.table("../data/FeatureoverlapPeaks/ShuffledPeak_overRepElements.txt", col.names = c("repCHR", "repStart", "repEnd", "Type", "length", "repStrand", "peakChr", "peakStart", "peakEnd", "peak", "score", "peakStrand"),stringsAsFactors = F)

repEl_shuf_sum= repEl_shuf %>% group_by(Type) %>% summarise(NperType_shuf=n()) %>% mutate(Proppertype_shuf=NperType_shuf/nrow(repEl_shuf)) 
```

Join both:  

```{r}
repEl_both=repEl_sum %>% full_join(repEl_shuf_sum, by="Type")

#fill NAs with 0  

repEl_both$Proppertype= repEl_both$Proppertype %>% replace_na(0)
repEl_both$NperType= repEl_both$NperType %>% replace_na(0)
repEl_both$Proppertype_shuf= repEl_both$Proppertype_shuf %>% replace_na(0)
repEl_both$NperType_shuf= repEl_both$NperType_shuf %>% replace_na(0)
```



Hyper geometric:  

```{r}
repEl_both_sig= repEl_both %>% mutate(Peak=sum(repEl_both$NperType), Peak_Shuf=sum(repEl_both$NperType_shuf), Chosen=NperType+NperType_shuf) %>% mutate(hyper=phyper(NperType, Peak, Peak_Shuf, Chosen, lower.tail = F)) %>% mutate(sig=ifelse(hyper<.05, "yes", "no")) %>% filter(sig=="yes") %>% arrange(desc(Proppertype))


repEl_both_sig

write.table(repEl_both_sig, "../data/FeatureoverlapPeaks/RepElementsWShuf.txt", col.names = T, row.names = F, quote=F)
```



