---
title: "Evaluate Brain Data"
author: "Briana Mittleman"
date: "7/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I downloaded the brain 3' seq data from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM747470 and I want to use this analysis to see how similar their peaks are to ours eventhough the data is from different cell types.  

First I will use the bedtools jaccard function to explore the overlaps. It will give me one stat that is the length(intersection)/length(union) - length(intersection). Here I can have file A brain peaks and file B be our peaks to see the similarity between the sets.  

```{bash, eval=F}
#!/bin/bash

#SBATCH --job-name=jaccard_brain
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=jacard_brain.out
#SBATCH --error=jacard_brain.err
#SBATCH --partition=broadwl
#SBATCH --mem=16G
#SBATCH --mail-type=END



module load Anaconda3
source activate three-prime-env 

bedtools jaccard -a /project2/gilad/briana/threeprimeseq/data/derti_brain/GSM747470_human_brain.sites.clustered.hg19.sort.bed -b /project2/gilad/briana/threeprimeseq/data/peaks/APApeaksYL.total.bed  > /project2/gilad/briana/threeprimeseq/data/derti_brain/total.jaccard.txt 
```


Results: 
intersection	union-intersection	jaccard	n_intersections
21371	25414133	0.00084091	21352

The brain set has 89110 peaks and our set has 288350. I will filter ours by score then see if the top 25% have a higher overlap percentage.  

```{r}
library(workflowr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
```


```{r}
YL_peaks=read.table("../data/bedgraph_peaks/APApeaks.bed", col.names = c("chr", "start", "end", "count", "strand", "score")) %>% mutate(length=end-start)
```

I want the counts for the top 25% of the peaks. 

```{r}
quantile(YL_peaks$count)
```

I will subset the peaks by having a count > 61.  

```{bash, eval=F}
awk '$4 >= 60 {print}' APApeaksYL.total.bed > APApeaksYL.top25.total.bed
```

I can rerun the jaccard with this and see if it changes, this new file has 72877 peaks.  

Results:   
intersection	union-intersection	jaccard	n_intersections
13221	6452066	0.00204911	13207


The proportion of overlap increased. Next I can try to make plots where I seperate my peaks by if they have a corresponding one in the brain file then plot the scores. To do this I will first use bedtool intersect to get just my peaks that contain a peak in the brain file. I can then use dplyr to merge them. 

Here A is my file and B is the brain file. 

```{bash, eval=F}
#!/bin/bash

#SBATCH --job-name=int_brain
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=int.brain.out
#SBATCH --error=int.brain.err
#SBATCH --partition=broadwl
#SBATCH --mem=16G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env 

bedtools intersect -wa -a /project2/gilad/briana/threeprimeseq/data/peaks/APApeaksYL.total.bed  -b /project2/gilad/briana/threeprimeseq/data/derti_brain/GSM747470_human_brain.sites.clustered.hg19.sort.bed > /project2/gilad/briana/threeprimeseq/data/derti_brain/APApeaksYL.total.inbrain.bed 


```
The resulting file has 21378 peaks.  

```{r}
YL_peaks_overlap=read.table("../data/APApeaksYL.total.inbrain.bed", col.names = c("chr", "start", "end", "count", "strand", "score")) %>% mutate(length=end-start) %>% mutate(in_brain="Y")
```


Now I need to join them.  

```{r}
YL_peaks_join=YL_peaks %>% full_join(YL_peaks_overlap, by = c("chr", "start", "end", "count", "strand", "score", "length"))

YL_peaks_join$in_brain[is.na(YL_peaks_join$in_brain)]="N"

YL_peaks_join_sel=YL_peaks_join %>% select(count, in_brain)
  
```

Plot these.  


```{r}
ggplot(YL_peaks_join_sel, aes(y=log10(count), x=in_brain, fill=in_brain)) + geom_boxplot() + labs(x="Peak called in brain dataset", y="log10 Score", title="Peak score distribution by inclusion in brain dataset")

```


```{r}
ggplot(YL_peaks_join_sel, aes(x=log10(count), fill=in_brain), bins=50) + geom_density(position="identity", alpha=.5) + labs(x="log10 of Score", title="Distribution of log10 Scores in peaks included in brain dataset")
```

It would be better if the background was just a random subset of the same number. There are 21378 included peaks so I should select a random 21378 to make a background distribution.  

```{r}
samp_YLpeaks= sample_n(YL_peaks, 21378)

ggplot() + geom_histogram(data=samp_YLpeaks, aes(log10(count)), bins=100) + geom_histogram(data=YL_peaks_overlap, aes(log10(count)),fill="Red", bins=100) + labs(x="Log10 of Score", title="Scores in Overlapping set compared to scores in random set")
```

