---
title: "Initial_gene_level"
author: "Briana Mittleman"
date: "5/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I will use this analysis to take a look at the initial protein conding gene counts.

```{r}
library(workflowr)
library(ggplot2)
library(dplyr)
library(edgeR)
```


Imput the data that was created from my coding gene rule in the snakefile.  

```{r}
N_18486_cov= read.table("../data/gene_cov/YL-SP-18486-N_S10_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))

T_18486_cov= read.table("../data/gene_cov/YL-SP-18486-T_S9_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18497_cov= read.table("../data/gene_cov/YL-SP-18497-N_S12_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18497_cov= read.table("../data/gene_cov/YL-SP-18497-T_S11_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18500_cov= read.table("../data/gene_cov/YL-SP-18500-N_S20_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18500_cov= read.table("../data/gene_cov/YL-SP-18500-T_S19_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18505_cov= read.table("../data/gene_cov/YL-SP-18505-N_S2_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18505_cov= read.table("../data/gene_cov/YL-SP-18505-T_S1_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18508_cov= read.table("../data/gene_cov/YL-SP-18508-N_S6_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18508_cov= read.table("../data/gene_cov/YL-SP-18508-T_S5_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18853_cov= read.table("../data/gene_cov/YL-SP-18853-N_S32_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18853_cov= read.table("../data/gene_cov/YL-SP-18853-T_S31_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_18870_cov= read.table("../data/gene_cov/YL-SP-18870-N_S24_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_18870_cov= read.table("../data/gene_cov/YL-SP-18870-T_S23_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19128_cov= read.table("../data/gene_cov/YL-SP-19128-N_S30_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19128_cov= read.table("../data/gene_cov/YL-SP-19128-T_S29_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19141_cov= read.table("../data/gene_cov/YL-SP-19141-N_S18_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19141_cov= read.table("../data/gene_cov/YL-SP-19141-T_S17_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19193_cov= read.table("../data/gene_cov/YL-SP-19193-N_S22_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19193_cov= read.table("../data/gene_cov/YL-SP-19193-T_S21_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19209_cov= read.table("../data/gene_cov/YL-SP-19209-N_S16_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19209_cov= read.table("../data/gene_cov/YL-SP-19209-T_S15_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19223_cov= read.table("../data/gene_cov/YL-SP-19223-N_S8_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19223_cov= read.table("../data/gene_cov/YL-SP-19233-T_S7_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19225_cov= read.table("../data/gene_cov/YL-SP-19225-N_S28_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19225_cov= read.table("../data/gene_cov/YL-SP-19225-T_S27_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19238_cov= read.table("../data/gene_cov/YL-SP-19238-N_S4_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19238_cov= read.table("../data/gene_cov/YL-SP-19238-T_S3_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19239_cov= read.table("../data/gene_cov/YL-SP-19239-N_S14_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19239_cov= read.table("../data/gene_cov/YL-SP-19239-T_S13_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
N_19257_cov= read.table("../data/gene_cov/YL-SP-19257-N_S26_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))
T_19257_cov= read.table("../data/gene_cov/YL-SP-19257-T_S25_R1_001-genecov.txt", stringsAsFactors = FALSE, header = F, col.names = c("chr", "start", "end", "gene", "score", "strand", "count" ))


```
Look at the total libraries first:  

```{r}
total_count_matrix=cbind(T_18486_cov$count, T_18497_cov$count, T_18500_cov$count, T_18505_cov$count, T_18508_cov$count, T_18853_cov$count, T_18870_cov$count, T_19128_cov$count, T_19141_cov$count, T_19193_cov$count, T_19209_cov$count, T_19223_cov$count, T_19225_cov$count, T_19238_cov$count,T_19239_cov$count, T_19257_cov$count)

#gene length vector
gene_length=T_18497_cov %>% mutate(genelength=end-start) %>% select(genelength) 
gene_length_vec=as.vector(gene_length$genelength)
total_count_matrix_cpm=cpm(total_count_matrix, log=T, gene.length=gene_length_vec )
```

Plot distribution of log2 cpm for total libraries.  

```{r}
plotDensities(total_count_matrix_cpm, legend = "bottomright", main="Pre-filtering total fraction")
```

Look at gene distributions for the nuclear fractions. 

```{r}
nuclear_count_matrix=cbind(N_18486_cov$count, N_18497_cov$count, N_18500_cov$count, N_18505_cov$count, N_18508_cov$count, N_18853_cov$count, N_18870_cov$count, N_19128_cov$count, N_19141_cov$count, N_19193_cov$count, N_19209_cov$count, N_19223_cov$count, N_19225_cov$count, N_19238_cov$count,N_19239_cov$count, N_19257_cov$count)

#cpm  

nuclear_count_matrix_cpm=cpm(nuclear_count_matrix, log=T, gene.length=gene_length_vec )

```


Plot distribution of log2 cpm for nuclear libraries.  

```{r}
plotDensities(nuclear_count_matrix_cpm, legend = "bottomright", main="Pre-filtering nuclear fraction")
```


The distributions look similar. I can filter based on alll of the libraries. I will filter for 1cpm in more than half of the libraries. After this I can ask how many genes are detected in each library.  

```{r}
all_count_matrix=cbind(T_18486_cov$count, T_18497_cov$count, T_18500_cov$count, T_18505_cov$count, T_18508_cov$count, T_18853_cov$count, T_18870_cov$count, T_19128_cov$count, T_19141_cov$count, T_19193_cov$count, T_19209_cov$count, T_19223_cov$count, T_19225_cov$count, T_19238_cov$count,T_19239_cov$count, T_19257_cov$count,N_18486_cov$count, N_18497_cov$count, N_18500_cov$count, N_18505_cov$count, N_18508_cov$count, N_18853_cov$count, N_18870_cov$count, N_19128_cov$count, N_19141_cov$count, N_19193_cov$count, N_19209_cov$count, N_19223_cov$count, N_19225_cov$count, N_19238_cov$count,N_19239_cov$count, N_19257_cov$count )


#cpm  

all_count_matrix_cpm=cpm(all_count_matrix, log=T, gene.length=gene_length_vec )
plotDensities(all_count_matrix_cpm, legend = "bottomright", main="Pre-filtering all libraries")
```

Filter:  

```{r}
keep.exprs=rowSums(all_count_matrix_cpm>1) >= 16
all_count_matrix_cpm_filt= all_count_matrix_cpm[keep.exprs,]

plotDensities(all_count_matrix_cpm_filt, legend = "bottomright", main="Post-filtering all libraries")
```

Post filtering we are left with 12461 protein coding genes.  
