---
title: "Investigate Peak to Gene Assignment"
author: "Briana Mittleman"
date: "12/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(workflowr)
library(cowplot)
library(tximport)
```


In looking at correlations and some examples, there is evidence the peak to gene assignment may be a problem. I am going to visualize the peaks in IGV. I will name them by the gene and look at them in the browser. 

The peak to gene annotations used in the feature counts to map reads back to the peaks is the following:  
* /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed.bed  

I need to change this a bit to have the name be the gene rather than the score:  

NamePeaksByGene.py
```{bash, eval=F}
#python  

CovnamedPeaks=open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed.bed", "r")
GeneNamedPeaks=open("/project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/PeaksNamedWithGeneAssignment.bed", "w")

for ln in CovnamedPeaks:
  chrom, start, end, num, cov, strand, transcript = ln.split()
  gene=transcript.split("-")[1]
  GeneNamedPeaks.write("%s\t%s\t%s\t%s\n"%(chrom,start,end,gene))

GeneNamedPeaks.close()

```


This was made based on the transcript annotation: ncbiRefSeq.mRNA.named.bed  

* /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named.bed  

The ends of the transcripts specfically are in:  

* /project2/gilad/briana/genome_anotation_data/ncbiRefSeq_endProtCodGenes_sort.txt


Ideas for Dilters:

* Cant be upstream of the gene, ex: chr2:135,558,075-135,604,343  

* maybe it cant be in another gene 

* we should include LINCs

* looks like we have a ton of low expressed intergenic peaks that should be filtered before we do the gene annotation


###Filter out intergenic peaks

As a first pass I want to filter out the peaks that are outside a gene body. While this may not be perfect it will help  alot with the intergenic noise.  

I need to overlap the named peaks with /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named.bed  and only keep the matches. I can use bedtools intersect.  

Rename the peaks according to convention to run an intesect.  

RenamePeaks4Intersect.py
```{bash,eval=F}
#python  
CovnamedPeaks=open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed.bed", "r")
GeneNamedPeaks=open("/project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_RENAMED.bed", "w")  

for ln in CovnamedPeaks:
  chrom, start, end, num, cov, strand, transcript = ln.split()
  gene=transcript.split("-")[1]
  start=int(start)
  end=int(end)
  GeneNamedPeaks.write("%s\t%d\t%d\t%s-%s\t%s\t%s\n"%(chrom,start,end,num,gene,cov,strand))

GeneNamedPeaks.close()


```

Remove CHR from the refseq annpotation:  
```{bash,eval=F}
sed 's/^chr//' /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named.bed > /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed
```



Filter4GenicPeaks.sh
```{bash,eval=F}

#!/bin/bash


#SBATCH --job-name=Filter4GenicPeaks
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=Filter4GenicPeaks.out
#SBATCH --error=Filter4GenicPeaks.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


bedtools intersect -wa -s -a /project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_RENAMED.bed -b /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed > /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies.bed
```

This is printing them multiple times.  

```{bash,eval=F}
uniq /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies.bed > /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodiesUNIQ.bed
```

Now I need to make this an SAF to run feature counts.  


bed2saf_peaksInGenicReg.py
 
 
```{bash,eval=F}
from misc_helper import *

fout = open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodiesUNIQ.SAF",'w')
fout.write("GeneID\tChr\tStart\tEnd\tStrand\n")
for ln in open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodiesUNIQ.bed"):
    chrom, start, end, name, score, strand = ln.split()
    namenum=name.split("-")[0]
    name_i=int(namenum)
    start_i=int(start)
    end_i=int(end)
    gene_only=name.split("-")[1]
    ID = "peak%d:%s:%d:%d:%s:%s"%(name_i, chrom, start_i, end_i, strand, gene_only)
    fout.write("%s\t%s\t%d\t%d\t%s\n"%(ID, chrom, start_i, end_i, strand))
fout.close()
```

Run Feature Counts  
PeaksinGenicRegion_fc_TN.sh  
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=PeaksinGenicRegion_fc_TN
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=PeaksinGenicRegion_fc_TN.out
#SBATCH --error=PeaksinGenicRegion_fc_TN.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodiesUNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Total.fc /project2/gilad/briana/threeprimeseq/data/sort/*-T-*-sort.bam -s 2

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodiesUNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Nuclear.fc /project2/gilad/briana/threeprimeseq/data/sort/*-N-*-sort.bam -s 2


```

Lastly I will need to fix the headers.  


fix_head_fc_genicPeak_tot.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Total.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Total_fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()
```

fix_head_fc_genicPeak_nuc.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Nuclear.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegion_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Nuclear_fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()
```


Pull these into R and look at the correlation between the sum of the peaks by gene and the transcripts counts from RNA seq.  

TPM counts from Kalisto
```{r}
tx2gene=read.table("../data/RNAkalisto/ncbiRefSeq.txn2gene.txt" ,header= F, sep="\t", stringsAsFactors = F)

txi.kallisto.tsv <- tximport("../data/RNAkalisto/abundance.tsv", type = "kallisto", tx2gene = tx2gene,countsFromAbundance="lengthScaledTPM" )
```
In previous analysis I did not account for gene length. Here I am going to standardize by length because I am taking a sum over a gene body.  

Import gene lengths:  
```{r}
geneLengthNames=c("CHR", "start", "end", "gene", "score", "strand")
geneLengths=read.table("../data/UnderstandPeaksQC/refseq.ProteinCoding.bed", header=F, stringsAsFactors = F, col.names = geneLengthNames) %>% mutate(length=end-start) %>% select(gene, length)
```


Look at the correlation with the total:  

I am using the sum of the counts in a gene divided by how many million reads mapped. I am also filtering out peaks with less than 10 reads in this individual.  
```{r}
total_Cov_18486=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% filter(X18486_T>0) %>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_T)) %>% mutate(GeneSumNorm=GeneSum/10.8) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}
TXN_abund=as.data.frame(txi.kallisto.tsv$abundance) %>% rownames_to_column(var="gene")
colnames(TXN_abund)=c("gene", "TPM")

TXN_NormGene=TXN_abund %>% inner_join(total_Cov_18486,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene=TXN_NormGene %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Tot=ggplot(TXN_NormGene, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Total", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.48") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Tot       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene)) 

cor.test(log10(TXN_NormGene$TPM),log10(TXN_NormGene$GeneSumSt))
```

Try this with nuclear  

```{r}
nuclear_Cov_18486=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_Genic.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% filter(X18486_N>10) %>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_N)) %>% mutate(GeneSumNorm=GeneSum/11.4) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}

TXN_NormGene_nuc=TXN_abund %>% inner_join(nuclear_Cov_18486,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_nuc=TXN_NormGene_nuc %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Nuc=ggplot(TXN_NormGene_nuc, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Nuclear", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.37") + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Nuc    
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_nuc)) 

cor.test(log10(TXN_NormGene_nuc$TPM),log10(TXN_NormGene_nuc$GeneSumSt))
```


This just said it had to be in a gene body not the specific gene body. This could be a problem still. For example in the SSPO locus chr7:149,521,993-149,543,749. Here the peaks are closer to the end of the SSPO but are in the gene body of the next gene downstream. 

Histones dont have a polyA tail- the HIST1H4C peak is most likely misprimming (chr6:26,102,306-26,110,443)  

Filter out overlapping genes:  


Count overlaps in origial file:  
```bedtools merge -i IN.bed -c 1 -o count > counted```

countGeneOverlap.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=countGeneOverlap
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=countGeneOverlap.out
#SBATCH --error=countGeneOverlap.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


bedtools merge -i /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR.bed -c 1 -o count > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap.bed  
```


Filter out these rows: 
```awk '/\t1$/{print}' counted > filtered```

```{bash,eval=F}
awk '/\t1$/{print}' /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap.bed > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap.filtered.bed
```

Intersect with original input to only keep the ones in both sets.  
```bedtools intersect -a IN.bed -b filtered -wa > OUT.bed```

findGeneswithoutOverlap.sh  
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=findGeneswithoutOverlap.sh
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=findGeneswithoutOverlap.out
#SBATCH --error=findGeneswithoutOverlap.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

bedtools intersect -a /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR.bed -b /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap.filtered.bed -wa > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.NonOverlapGenes.bed 

```


Finally overlap with the mRNA file to only keep the transcripts in these genes. This may be easiest in python /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed


subsetmRNAforNonOverlapGenes.py
```{bash,eval=F}
geneFile=open("/project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.NonOverlapGenes.bed", "r") 

mRNAFile=open("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed", "r")

outFile=open("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR_NoneOverlapingGenes.bed", "w")  

#make list of non overlapping genes  

keep=[]
for ln in geneFile:
  keep.append(ln.split()[3])

for ln in mRNAFile: 
  if ln.split()[4] in keep:
    outFile.write(ln)
outFile.close()

```
Filter peaks on this resutls  

Filter4GenicPeaks_noOverlap.sh
```{bash,eval=F}

#!/bin/bash


#SBATCH --job-name=Filter4GenicPeaks
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=Filter4GenicPeaks.out
#SBATCH --error=Filter4GenicPeaks.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


bedtools intersect -wa -s -a /project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_RENAMED.bed -b /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR_NoneOverlapingGenes.bed> /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap.bed
```

This is printing them multiple times.  

```{bash,eval=F}
uniq /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap.bed > /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap_UNIQ.bed
```


Make this an SAF to run FC  


bed2saf_peaksInGenicReg_noOVERLAP.py
 
 
```{bash,eval=F}
from misc_helper import *

fout = open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap_UNIQ.SAF",'w')
fout.write("GeneID\tChr\tStart\tEnd\tStrand\n")
for ln in open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap_UNIQ.bed"):
    chrom, start, end, name, score, strand = ln.split()
    namenum=name.split("-")[0]
    name_i=int(namenum)
    start_i=int(start)
    end_i=int(end)
    gene_only=name.split("-")[1]
    ID = "peak%d:%s:%d:%d:%s:%s"%(name_i, chrom, start_i, end_i, strand, gene_only)
    fout.write("%s\t%s\t%d\t%d\t%s\n"%(ID, chrom, start_i, end_i, strand))
fout.close()
```


Run Feature Counts  
PeaksinGenicRegion_NoneOverlapGenes_fc_TN.sh  
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=PeaksinGenicRegion_NoneOverlapGenes_fc_TN
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=PeaksinGenicRegion_NoneOverlapGenes_fc_TN.out
#SBATCH --error=PeaksinGenicRegion_NoneOverlapGenes_fc_TN.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap_UNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Total.fc /project2/gilad/briana/threeprimeseq/data/sort/*-T-*-sort.bam -s 2

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlap_UNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Nuclear.fc /project2/gilad/briana/threeprimeseq/data/sort/*-N-*-sort.bam -s 2


```

Lastly I will need to fix the headers.  


fix_head_fc_genicPeakNoOverlap_tot.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Total.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Total_fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()
```

fix_head_fc_genicPeakNoOverlap_nuc.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Nuclear.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlap_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Nuclear_fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()

```


Pull these onto my computer:  

Use no filter and standardization scheme:  

```{r}
total_Cov_18486_noOver=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T)%>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_T)) %>% mutate(GeneSumNorm=GeneSum/10.8) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}

TXN_NormGene_noOverlap=TXN_abund %>% inner_join(total_Cov_18486_noOver,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_noOverlap=TXN_NormGene_noOverlap %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Tot_noOver=ggplot(TXN_NormGene_noOverlap, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Total", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.49") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Tot_noOver       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlap)) 

cor.test(log10(TXN_NormGene_noOverlap$TPM),log10(TXN_NormGene_noOverlap$GeneSumSt))
```

```{r}
nuclear_Cov_18486_noOver=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_N)) %>% mutate(GeneSumNorm=GeneSum/11.4) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}

TXN_NormGene_noOverlap_nuc=TXN_abund %>% inner_join(nuclear_Cov_18486_noOver,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_noOverlap_nuc=TXN_NormGene_noOverlap_nuc %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Nuc_noOver=ggplot(TXN_NormGene_noOverlap_nuc, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Nuclear", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.5") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Nuc_noOver       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlap_nuc)) 

cor.test(log10(TXN_NormGene_noOverlap_nuc$TPM),log10(TXN_NormGene_noOverlap_nuc$GeneSumSt),method="spearman")
```

It looks like a more stringent RNA seq filter could help. Lets say  log(TPM)>-1.5  
```{r}
TXN_abund_filt=TXN_abund %>% filter(log(TPM) > -1.5)
TXN_NormGene_noOverlap_filt=TXN_abund_filt %>% inner_join(total_Cov_18486_noOver,by="gene")
TXN_NormGene_noOverlap_filt=TXN_NormGene_noOverlap_filt %>% filter(TPM >0) %>% filter(GeneSumSt>0)
corr_18486Tot_noOver_filt=ggplot(TXN_NormGene_noOverlap_filt, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Total filtered", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.49") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Tot_noOver_filt      
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlap_filt)) 
cor.test(log10(TXN_NormGene_noOverlap_filt$TPM),log10(TXN_NormGene_noOverlap_filt$GeneSumSt),method="spearman")

```



This does not take care of genes that are near each other such as what is going on with SSPO


##Filter genes that are too close together

This is similar to the overlap problem but I want to extend the genes.  

I can a python script to subtract 2000 bases from the start and add 2000 to the end 


*/project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR.bed
 
 
extendGenes.py
```{bash,eval=F}
inFile=open("/project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR.bed", "r")

outFile=open("/project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR_EXT.bed", "w")\


for ln in inFile:
  chrom, start, end, gene, score, strand = ln.split()
  start_ex=int(start)-2000
  if start_ex <0: 
    start_ex=0
  end_ex=int(end)+2000
  outFile.write("%s\t%d\t%d\t%s\t%s\t%s\n"%(chrom, start_ex, end_ex, gene, score, strand))
  
outFile.close()
```


Now I can run the merge:  

countGeneOverlap_EXT.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=countGeneOverlap_EXT
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=countGeneOverlap_EXT.out
#SBATCH --error=countGeneOverlap_EXT.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


bedtools merge -i /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR_EXT.bed -c 1 -o count > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap_EXT.bed  
```




Filter out these rows: 
```awk '/\t1$/{print}' counted > filtered```

```{bash,eval=F}
awk '/\t1$/{print}' /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap_EXT.bed > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap_EXT_filter.bed
```


Intersect with original input to only keep the ones in both sets.  
```bedtools intersect -a IN.bed -b filtered -wa > OUT.bed```

findGeneswithoutOverlap_EXT.sh  
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=findGeneswithoutOverlap_EXT.sh
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=findGeneswithoutOverlap_EXT.out
#SBATCH --error=findGeneswithoutOverlap_EXT.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

bedtools intersect -a /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.noCHR.bed -b /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.countGeneOverlap_EXT_filter.bed -wa > /project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.NonOverlapGenes_EXT.bed 

```
Finally overlap with the mRNA file to only keep the transcripts in these genes. This may be easiest in python /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed


subsetmRNAforNonOverlapGenes_EXT.py
```{bash,eval=F}
geneFile=open("/project2/gilad/briana/genome_anotation_data/refseq.ProteinCoding.NonOverlapGenes_EXT.bed", "r") 

mRNAFile=open("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR.bed", "r")

outFile=open("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR_NoneOverlapingGenes_EXT.bed", "w")  

#make list of non overlapping genes  

keep=[]
for ln in geneFile:
  keep.append(ln.split()[3])

for ln in mRNAFile: 
  if ln.split()[4] in keep:
    outFile.write(ln)
outFile.close()

```
Filter4GenicPeaks_noOverlap_EXT.sh
```{bash,eval=F}

#!/bin/bash


#SBATCH --job-name=Filter4GenicPeaks_EXT
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=Filter4GenicPeaks_EXT.out
#SBATCH --error=Filter4GenicPeaks_EXT.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env


bedtools intersect -wa -s -a /project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_RENAMED.bed -b /project2/gilad/briana/genome_anotation_data/ncbiRefSeq.mRNA.named_noCHR_NoneOverlapingGenes_EXT.bed > /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT.bed
```

This is printing them multiple times.  

```{bash,eval=F}
uniq /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT.bed > /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT_UNIQ.bed
```
Make this an SAF to run FC  


bed2saf_peaksInGenicReg_noOVERLAPEXT.py
 
 
```{bash,eval=F}
from misc_helper import *

fout = open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT_UNIQ.SAF",'w')
fout.write("GeneID\tChr\tStart\tEnd\tStrand\n")
for ln in open("/project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT_UNIQ.bed"):
    chrom, start, end, name, score, strand = ln.split()
    namenum=name.split("-")[0]
    name_i=int(namenum)
    start_i=int(start)
    end_i=int(end)
    gene_only=name.split("-")[1]
    ID = "peak%d:%s:%d:%d:%s:%s"%(name_i, chrom, start_i, end_i, strand, gene_only)
    fout.write("%s\t%s\t%d\t%d\t%s\n"%(ID, chrom, start_i, end_i, strand))
fout.close()
```


Run Feature Counts  
PeaksinGenicRegion_NoneOverlapGenesEXT_fc_TN.sh  
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=PeaksinGenicRegion_NoneOverlapGenesEXT_fc_TN
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=PeaksinGenicRegion_NoneOverlapGenesEXT_fc_TN.out
#SBATCH --error=PeaksinGenicRegion_NoneOverlapGenesEXT_fc_TN.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT_UNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Total.fc /project2/gilad/briana/threeprimeseq/data/sort/*-T-*-sort.bam -s 2

featureCounts -O -a /project2/gilad/briana/threeprimeseq/data/mergedPeaks_comb_inGeneBody/filtered_APApeaks_merged_allchrom_refseqTrans.noties_sm.fixed_inGeneBodies_noGeneOverlapEXT_UNIQ.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Nuclear.fc /project2/gilad/briana/threeprimeseq/data/sort/*-N-*-sort.bam -s 2


```
Lastly I will need to fix the headers.  


fix_head_fc_genicPeakNoOverlapEXT_tot.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Total.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Total_fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()
```

fix_head_fc_genicPeakNoOverlapEXT_nuc.py
```{bash,eval=F}
infile= open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Nuclear.fc", "r")
fout = open("/project2/gilad/briana/threeprimeseq/data/PeakInGenecRegionNoOverlapEXT_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Nuclear_Fixed.fc",'w')
for line, i in enumerate(infile):
    if line == 1:
        i_list=i.split()
        libraries=i_list[:6]
        for sample in i_list[6:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n')
    else :
        fout.write(i)
fout.close()

```

Pull these onto my computer:  

Use no filter and standardization scheme:  

```{r}
total_Cov_18486_noOverEXT=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T)%>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_T)) %>% mutate(GeneSumNorm=GeneSum/10.8) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}

TXN_NormGene_noOverlapEXT=TXN_abund %>% inner_join(total_Cov_18486_noOverEXT,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_noOverlapEXT=TXN_NormGene_noOverlapEXT %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Tot_noOverEXT=ggplot(TXN_NormGene_noOverlapEXT, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Total", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.46") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Tot_noOverEXT       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlapEXT)) 

cor.test(log10(TXN_NormGene_noOverlapEXT$TPM),log10(TXN_NormGene_noOverlapEXT$GeneSumSt),method="spearman")
```
In this test, I was looking at 5871 genes. 

```{r}
nuclear_Cov_18486_noOverEXT=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlapEXT.Nuclear_Fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_N)) %>% mutate(GeneSumNorm=GeneSum/11.4) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```

Join the data frames.  

```{r}

TXN_NormGene_noOverlap_nucEXT=TXN_abund %>% inner_join(nuclear_Cov_18486_noOverEXT,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_noOverlap_nucEXT=TXN_NormGene_noOverlap_nucEXT %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18486Nuc_noOverEXT=ggplot(TXN_NormGene_noOverlap_nucEXT, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Nuclear", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.5") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18486Nuc_noOverEXT       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlap_nucEXT)) 

cor.test(log10(TXN_NormGene_noOverlap_nucEXT$TPM),log10(TXN_NormGene_noOverlap_nucEXT$GeneSumSt),method="spearman")
```

In this test, I was looking at 6475 genes. 

##Check another line  


```{r}
total_Cov_18497_noOver=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant_GenicNoOverlap.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:8] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18497_T)%>%  group_by(gene) %>% summarize(GeneSum=sum(X18497_T)) %>% mutate(GeneSumNorm=GeneSum/10.8) %>% inner_join(geneLengths, by="gene") %>% mutate(GeneSumSt=GeneSum/length)

```



Join the data frames.  

```{r}

TXN_NormGene_noOverlap_497=TXN_abund %>% inner_join(total_Cov_18497_noOver,by="gene")
```


Remove rows with 0 counts and Plot:  

```{r}
TXN_NormGene_noOverlap_497=TXN_NormGene_noOverlap_497 %>% filter(TPM>0) %>% filter(GeneSumSt>0)
corr_18497Tot_noOver=ggplot(TXN_NormGene_noOverlap_497, aes(x=log10(TPM), y= log10(GeneSumSt))) + geom_point() + labs(title="Total", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM),y=log10(GeneSumSt)),method = "lm") + annotate("text",x=3, y=5,label="R2=.49") +geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

#+ geom_text(aes(label=gene),hjust=0, vjust=0)
       
corr_18497Tot_noOver       
summary(lm(log10(TPM)~log10(GeneSumSt),TXN_NormGene_noOverlap_497)) 

cor.test(log10(TXN_NormGene_noOverlap_497$TPM),log10(TXN_NormGene_noOverlap_497$GeneSumSt),method="spearman")
```

This shows me the effect is not line specific because the correlation is the same when i use the RNA from one line and 3' from another.  



