---
title: "Nuclear vs Total nominal associations"
author: "Briana Mittleman"
date: "3/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Look for the snps that are QTLs in the nuclear fraction but not the total fraction. To do this I want to plot the association in the total fraction vs. the association in the nuclear fraction. I can look for snps that fall off the 1:1 line towards the nuclear fraction.  

I need to look for the snps tested in both fractions. The nominal results are in:

* /project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt   
* /project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt  

This will not include the 18982867 more assocaitions in nucelar. I can only look at those that are tested in both analyses. 

Format of file: 
* peakID
* snp
* dist 
* pval
* slope (effect size)  

I can make a dictionary with gene:snp:peak as the keys and a dictionary for the values- the inner dictionary will have the fraction as the key and the pvalue as the value  



totalandnuclear_commonassociation.py
```{bash,eval=F}
totRes=open("/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt","r")
nucRes=open("/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt", "r")  


outfile=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt", "w")

resDict={}
for ln in totRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    snp=ln.split()[1]
    id=gene + ":" + peak + ":" + snp
    pval=ln.split()[3]
    resDict[id]={}
    resDict[id]["Total"]=pval

for ln in nucRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    snp=ln.split()[1]
    id=gene + ":" + peak + ":" + snp
    pval=ln.split()[3]
    if id in resDict.keys():
        resDict[id]["Nuclear"]=pval
    else:
        continue   
        
#now i have a double dictionary. i need to write it out. i want id, the total pval, then the nuc pval  
for outer, inner in resDict.items():
    id=outer
    outPval=[]
    for key in inner:
        outPval.append(inner[key])
    print(outPval)
    if(len(outPval))==2:
         outfile.write("%s\t%s\t%s\n"%(id, outPval[0], outPval[1]))


outfile.close()        
        

```


run it: 
run_totalandnuclear_commonassociation.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_totalandnuclear_commonassociation
#SBATCH --account=pi-yangili1
#SBATCH --time=5:00:00
#SBATCH --output=run_totalandnuclear_commonassociation.out
#SBATCH --error=run_totalandnuclear_commonassociation.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env


python totalandnuclear_commonassociation.py

```

Tot assoc:     105456802
nuclear assoc: 124439669
common:        94176434
This means 41 percent are common. This means in the make phenotpye step, different genes and peaks are removed. 
Results: /project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt  


makeTotalNucAssocPlot.R
```{r, eval=F}
library(tidyverse)
library(data.table)

file=fread("../data/NucSpecQTL/TotNucRes_overlapassociations.txt", col.names =c("ID","Total", "Nuclear"))

cor_plot=ggplot(file, aes(x=-log10(Total), y=-log10(Nuclear))) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + labs(title="Common Peak/Snp/ID associations Total and Nuclear", x="-log10(Total Pval)", y="-log10(Nuclear Pval)") + geom_smooth(aes(x=-log10(Total),y=-log10(Nuclear)),method = "lm")


#print(summary(lm(data=file, -log10(Total) ~ -log10(Nuclear))))

#cor_plot

#ggsave(plot, "/project2/gilad/briana/threeprimeseq/output/TotalvNucPavalCommonAssoc.png")  



```
in python with pyplotlib:


```{bash,eval=F}

import matplotlib.pyplot as plt
import numpy as np


inF=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt", "r")


total=[]
nuclear=[]
for ln in inF:
  loca, tot, nuc = ln.split()
  total.append(float(tot))
  nuclear.append(float(nuc))

  
np.asarray(total)
np.asarray(nuclear)
diff =np.subtract(total, nuclear)

plt.plot(total,nuclear)
plt.savefig("/project2/gilad/briana/threeprimeseq/output/TotalvNucPavalCommonAssoc.png")
```



run_makeTotalNucAssocPlot.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_makeTotalNucAssocPlot
#SBATCH --account=pi-yangili1
#SBATCH --time=5:00:00
#SBATCH --output=run_makeTotalNucAssocPlot.out
#SBATCH --error=run_makeTotalNucAssocPlot.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env

Rscript makeTotalNucAssocPlot.R
```


These are too many values: I need to use the permuted. this will jsut be the pvalue for the same peak:gene combo  

##try with perm
totalandnuclear_commonassociationPerm.py
```{bash,eval=F}
totRes=open("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno.NoMP_sm_quant.Total.fixed.pheno_5perc_permRes.txt","r")
nucRes=open("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno.NoMP_sm_quant.Nuclear.fixed.pheno_5perc_permRes.txt", "r")  


outfile=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_Permoverlapassociations.txt", "w")

resDict={}
for ln in totRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    id=gene + ":" + peak 
    pval=ln.split()[-2]
    resDict[id]={}
    resDict[id]["Total"]=pval

for ln in nucRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    id=gene + ":" + peak
    pval=ln.split()[-2]
    if id in resDict.keys():
        resDict[id]["Nuclear"]=pval
    else:
        continue   
        
#now i have a double dictionary. i need to write it out. i want id, the total pval, then the nuc pval  
for outer, inner in resDict.items():
    id=outer
    outPval=[]
    for key in inner:
        outPval.append(inner[key])
    print(outPval)
    if(len(outPval))==2:
         outfile.write("%s\t%s\t%s\n"%(id, outPval[0], outPval[1]))


outfile.close()        
        

```


```{r}
library(tidyverse)
library(data.table)

file=fread("../data/NucSpecQTL/TotNucRes_Permoverlapassociations.txt", col.names =c("ID","Total", "Nuclear"))

cor_plot=ggplot(file, aes(x=Total, y=Nuclear)) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + labs(title="Common Peak associations Total and Nuclear") 

cor_plot

```
This is similar to what was happening in the nominal case.  


I think I need to plot just the total in nuclaer and nuclear in total. this way i can see the outliers. 
