---
title: "Nuclear vs Total nominal associations"
author: "Briana Mittleman"
date: "3/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Look for the snps that are QTLs in the nuclear fraction but not the total fraction. To do this I want to plot the association in the total fraction vs. the association in the nuclear fraction. I can look for snps that fall off the 1:1 line towards the nuclear fraction.  

I need to look for the snps tested in both fractions. The nominal results are in:

* /project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt   
* /project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt  

This will not include the 18982867 more assocaitions in nucelar. I can only look at those that are tested in both analyses. 

Format of file: 
* peakID
* snp
* dist 
* pval
* slope (effect size)  

I can make a dictionary with gene:snp:peak as the keys and a dictionary for the values- the inner dictionary will have the fraction as the key and the pvalue as the value  



totalandnuclear_commonassociation.py
```{bash,eval=F}
totRes=open("/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt","r")
nucRes=open("/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt", "r")  


outfile=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt", "w")

resDict={}
for ln in totRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    snp=ln.split()[1]
    id=gene + ":" + peak + ":" + snp
    pval=ln.split()[3]
    resDict[id]={}
    resDict[id]["Total"]=pval

for ln in nucRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    snp=ln.split()[1]
    id=gene + ":" + peak + ":" + snp
    pval=ln.split()[3]
    if id in resDict.keys():
        resDict[id]["Nuclear"]=pval
    else:
        continue   
        
#now i have a double dictionary. i need to write it out. i want id, the total pval, then the nuc pval  
for outer, inner in resDict.items():
    id=outer
    outPval=[]
    for key in inner:
        outPval.append(inner[key])
    print(outPval)
    if(len(outPval))==2:
         outfile.write("%s\t%s\t%s\n"%(id, outPval[0], outPval[1]))


outfile.close()        
        

```


run it: 
run_totalandnuclear_commonassociation.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_totalandnuclear_commonassociation
#SBATCH --account=pi-yangili1
#SBATCH --time=5:00:00
#SBATCH --output=run_totalandnuclear_commonassociation.out
#SBATCH --error=run_totalandnuclear_commonassociation.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env


python totalandnuclear_commonassociation.py

```

Tot assoc:     105456802
nuclear assoc: 124439669
common:        94176434
This means 41 percent are common. This means in the make phenotpye step, different genes and peaks are removed. 
Results: /project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt  


makeTotalNucAssocPlot.R
```{r, eval=F}
library(tidyverse)
library(data.table)

file=fread("../data/NucSpecQTL/TotNucRes_overlapassociations.txt", col.names =c("ID","Total", "Nuclear"))

cor_plot=ggplot(file, aes(x=-log10(Total), y=-log10(Nuclear))) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + labs(title="Common Peak/Snp/ID associations Total and Nuclear", x="-log10(Total Pval)", y="-log10(Nuclear Pval)") + geom_smooth(aes(x=-log10(Total),y=-log10(Nuclear)),method = "lm")


#print(summary(lm(data=file, -log10(Total) ~ -log10(Nuclear))))

#cor_plot

#ggsave(plot, "/project2/gilad/briana/threeprimeseq/output/TotalvNucPavalCommonAssoc.png")  



```
in python with pyplotlib:


```{bash,eval=F}

import matplotlib.pyplot as plt
import numpy as np


inF=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_overlapassociations.txt", "r")


total=[]
nuclear=[]
for ln in inF:
  loca, tot, nuc = ln.split()
  total.append(float(tot))
  nuclear.append(float(nuc))

  
np.asarray(total)
np.asarray(nuclear)
diff =np.subtract(total, nuclear)

plt.plot(total,nuclear)
plt.savefig("/project2/gilad/briana/threeprimeseq/output/TotalvNucPavalCommonAssoc.png")
```



run_makeTotalNucAssocPlot.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_makeTotalNucAssocPlot
#SBATCH --account=pi-yangili1
#SBATCH --time=5:00:00
#SBATCH --output=run_makeTotalNucAssocPlot.out
#SBATCH --error=run_makeTotalNucAssocPlot.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env

Rscript makeTotalNucAssocPlot.R
```


These are too many values: I need to use the permuted. this will jsut be the pvalue for the same peak:gene combo  

##try with perm
totalandnuclear_commonassociationPerm.py
```{bash,eval=F}
totRes=open("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno.NoMP_sm_quant.Total.fixed.pheno_5perc_permRes.txt","r")
nucRes=open("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno.NoMP_sm_quant.Nuclear.fixed.pheno_5perc_permRes.txt", "r")  


outfile=open("/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotNucRes_Permoverlapassociations.txt", "w")

resDict={}
for ln in totRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    id=gene + ":" + peak 
    pval=ln.split()[-2]
    resDict[id]={}
    resDict[id]["Total"]=pval

for ln in nucRes:
    gene=ln.split()[0].split(":")[-1].split("_")[0]
    peak=ln.split()[0].split(":")[-1].split("_")[-1]
    id=gene + ":" + peak
    pval=ln.split()[-2]
    if id in resDict.keys():
        resDict[id]["Nuclear"]=pval
    else:
        continue   
        
#now i have a double dictionary. i need to write it out. i want id, the total pval, then the nuc pval  
for outer, inner in resDict.items():
    id=outer
    outPval=[]
    for key in inner:
        outPval.append(inner[key])
    print(outPval)
    if(len(outPval))==2:
         outfile.write("%s\t%s\t%s\n"%(id, outPval[0], outPval[1]))


outfile.close()        
        

```


```{r}
library(tidyverse)
library(data.table)
library(cowplot)

file=fread("../data/NucSpecQTL/TotNucRes_Permoverlapassociations.txt", col.names =c("ID","Total", "Nuclear"))

cor_plot=ggplot(file, aes(x=Total, y=Nuclear)) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + labs(title="Common Peak associations Total and Nuclear") 

cor_plot

```
This is similar to what was happening in the nominal case.  


I think I need to plot just the total in nuclaer and nuclear in total. this way i can see the outliers. 


```{r}
names=c("SNP", "peak", "pval")
NucQTLinTot=read.table("../data/QTL_overlap/NucQTLs_inTotFractionRes.txt", stringsAsFactors = F, col.names = names)
TotQTLinNuc=read.table("../data/QTL_overlap/TotQTLs_inNucFractionRes.txt", stringsAsFactors = F, col.names = names)

QTL_names= c("pid", "nvar", "shape1", "shape2", "dummy", "SNP", "dist", "npval", "slope", "ppval", "bpval", "bh")
NucQTL=read.table("../data/ApaQTLs/NuclearapaQTLs.GeneLocAnno.noMP.5perc.10FDR.txt", col.names = QTL_names, stringsAsFactors = F) %>% separate(pid, into=c("chr", "start","end", "peakID"), sep=":") %>% separate(peakID, into=c("gene", "strand", "peak"), sep="_") %>% select(SNP, peak, ppval)

TotQTL=read.table("../data/ApaQTLs/TotalapaQTLs.GeneLocAnno.noMP.5perc.10FDR.txt", col.names = QTL_names, stringsAsFactors = F) %>% separate(pid, into=c("chr", "start","end", "peakID"), sep=":") %>% separate(peakID, into=c("gene", "strand", "peak"), sep="_") %>% select(SNP, peak, ppval)
```


Join these:

```{r}
TotQTL_bothfrac= TotQTL %>% inner_join(TotQTLinNuc, by=c("SNP", "peak"))
colnames(TotQTL_bothfrac)= c("SNP", "peak", "Total", "Nuclear")
```


```{r}
NucQTL_bothfrac= NucQTL %>% inner_join(NucQTLinTot, by=c("SNP", "peak"))
colnames(NucQTL_bothfrac)= c("SNP", "peak", "Nuclear", "Total")
```


All qtls together:

```{r}
allQTL=TotQTL_bothfrac %>% bind_rows(NucQTL_bothfrac)
```


```{r}
ggplot(allQTL, aes(x=Nuclear, y=Total)) + geom_point()
```

The pvalues are not consistent. I need to get the original pvalues from the nominal files.  


nominal associations for each QTL:  

getNominal4QTLs.py

```{bash,eval=F}

nucQTLs="/project2/gilad/briana/threeprimeseq/data/ApaQTLs/Nuclear.apaQTLs.sort.bed"  
totQTLs="/project2/gilad/briana/threeprimeseq/data/ApaQTLs/Total.apaQTLs.sort.bed"  

nucNom="/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Nuclear.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt"
totNom="/project2/gilad/briana/threeprimeseq/data/nominal_APAqtl_GeneLocAnno_noMP_5percUs/filtered_APApeaks_merged_allchrom_refseqGenes.GeneLocAnno_NoMP_sm_quant.Total.fixed.pheno_5perc.fc.gz.qqnorm_allNomRes.txt"

outtot="/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/TotQTLs_inTotFractionRes.txt"
outnuc="/project2/gilad/briana/threeprimeseq/data/NucSpecQTL/NucQTLs_inNucFractionRes.txt"


def sameFract(inRes, inQTL, out):
    fout=open(out, "w")
    qtl_dic={}
    #SNP is key, peak is value
    for ln in open(inQTL,"r"):
        snp=ln.split()[2]
        chrom=ln.split()[0]
        peak=ln.split()[3].split(":")[0]
        qtl=str(chrom) + ":" + str(snp)
        if qtl not in qtl_dic.keys():
            qtl_dic[qtl]=[peak]
        else: 
            qtl_dic[qtl].append(peak)
    #print(qtl_dic)
    for ln in open(inRes, "r"):
        pval=ln.split()[3]
        snp=ln.split()[1]
        peak=ln.split()[0].split(":")[3].split("_")[-1]
        if snp in qtl_dic.keys():
            if peak in qtl_dic[snp]:
                fout.write("%s\t%s\t%s\n"%(snp, peak, pval))   
    fout.close()
    

sameFract(nucNom, nucQTLs,outnuc)  
sameFract(totNom, totQTLs, outtot)  
                
```


run_getNominal4QTLs.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_getNominal4QTLs
#SBATCH --account=pi-yangili1
#SBATCH --time=5:00:00
#SBATCH --output=run_getNominal4QTLs.out
#SBATCH --error=run_getNominal4QTLs.err
#SBATCH --partition=bigmem2
#SBATCH --mem=100G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env

python getNominal4QTLs.py
```


I now want to plot the total vs nuclear nominal pvalues for the nuclear QTLs.  

```{r}
colnames(NucQTLinTot)=c("SNP", "peak", "totalPval")

NucQTLinNuc=read.table("../data/NucSpecQTL/NucQTLs_inNucFractionRes.txt", col.names = c("SNP", "peak","nuclearPval"), stringsAsFactors = F)


NucQTL_all= NucQTLinTot %>% inner_join(NucQTLinNuc, by=c("SNP", "peak"))

summary(lm(data=NucQTL_all,-log10(totalPval) ~ -log10(nuclearPval)))
```


```{r}
bothPval_nucqtl=ggplot(NucQTL_all, aes(x=-log10(totalPval), y=-log10(nuclearPval))) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + geom_smooth(aes(x=-log10(totalPval),y=-log10(nuclearPval)),method = "lm") + labs(title="Nominal Pvalues for Nuclear APA qtls")
#+geom_text(aes(label=SNP),hjust=0, vjust=0)

bothPval_nucqtl



```


```{r}
ggplot(NucQTL_all, aes(x=totalPval, y=nuclearPval)) + geom_point() + labs(title="Nominal Pval results for Nuclear QTLs") + geom_density2d(na.rm = TRUE, size = 1, colour = 'red')  +geom_vline(xintercept=.05)

nNucQTLNotSigTot= NucQTL_all %>% filter(totalPval>.05) %>% nrow()
nNucQTLNotSigTot
```

For Total QTLs:  

```{r}
colnames(TotQTLinNuc)=c("SNP", "peak", "nuclearPval")

TotQTLinTot=read.table("../data/NucSpecQTL/TotQTLs_inTotFractionRes.txt", col.names = c("SNP", "peak","totalPval"), stringsAsFactors = F)


TotQTL_all= TotQTLinNuc %>% inner_join(TotQTLinTot, by=c("SNP", "peak"))

summary(lm(data=NucQTL_all,-log10(nuclearPval) ~ -log10(totalPval)))
```



```{r}
bothPval_totqtl=ggplot(TotQTL_all, aes(x=-log10(nuclearPval), y=-log10(totalPval))) + geom_point() + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') + geom_smooth(aes(x=-log10(nuclearPval),y=-log10(totalPval)),method = "lm") + labs(title="Nominal Pvalues for Total APA qtls")
#+geom_text(aes(label=SNP),hjust=0, vjust=0)

bothPval_totqtl
```


```{r}
bothFracPlots=plot_grid(bothPval_nucqtl,bothPval_totqtl)
bothFracPlots

ggsave(bothFracPlots, file="../output/plots/NominalPvalcondQTL.png", height=7, width=12 )
```

```{r}
ggplot(TotQTL_all, aes(x=nuclearPval, y=totalPval)) + geom_point() + labs(title="Nominal Pval results for Total QTLs") + geom_density2d(na.rm = TRUE, size = 1, colour = 'red') 

nTotQTLNotSigNuc= TotQTL_all %>% filter(nuclearPval>.05) %>% nrow()
nTotQTLNotSigNuc
```

Is the difference significant. More likely to have nuclear specific qlts?

```{r}
phyper(118, 495, 272, 157, lower.tail = F)
```



Plot for this:

```{r}
compareSpecQTL=as.data.frame(cbind(fraction=c("Total QTL", "Total QTL","Nuclear QTL","Nuclear QTL"), SigInOther=c("No","Yes","No","Yes"),Value=c(.14,.86 ,.24,.76)))

compareSpecQTL$Value=as.numeric(as.character(compareSpecQTL$Value))


```

```{r}
qtls_oppfrac=ggplot(compareSpecQTL,aes(x=fraction, fill=SigInOther, y=Value))  + geom_bar(stat="identity") +scale_fill_manual(values=c("red","grey")) + labs(y="Proportion", title="QTLs by nominal significance in other fraction")  + annotate("text", label="p < .0005", y=1.2, x=1.5) + geom_segment(aes(x=1.1, xend=1.9, y=1.1, yend=1.1))

qtls_oppfrac

ggsave(qtls_oppfrac, file="../output/plots/QTLsSiginOppFraction.png")
```

