---
title: "Overlap molQTLs, Opposite Direction"
author: "Briana Mittleman"
date: "10/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the [OverlapMolQTL](overlapMolQTL.html) analysis I looked at significant molecular QTLs and asked if they are also significant snp:gene pairs in the ApaQTLs. In this analysis, I will look at the significant ApaQTLs and ask if the snp:gene pairs are significant in the other molecular phenotypes. I expect enrichment of low pvalues in protQTLs but less in RNA.  

I am going to complete this analysis first for the totalAPA QTLs.  


```{r}
library(workflowr)
library(reshape2)
library(tidyverse)
library(VennDiagram)
library(data.table)
library(qvalue)
```




##MolQTL pvalues for Total ApaQTLs  

sigTotAPAinMolPheno.R
```{r,eval=F}
#!/bin/rscripts

#this script creates takes in the permuted APAQTL results for the total fraction and nominal pvalues from the molecular phenotpye  molecular phenotype 

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(optparse)

geneNames=read.table("/project2/gilad/briana/genome_anotation_data/ensemble_to_genename.txt", sep="\t", header=T, stringsAsFactors = F)

tot_perm=read.table("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Total_transcript_permResBH.txt", header = T,stringsAsFactors=F)

sigSNPgene=tot_perm %>% filter(-log10(bh)>1) %>% separate(pid, into=c("chr", "start", "end", "id"), sep=":") %>% separate(id, into=c("Gene.name", "strand", "peaknum"), sep="_") %>% dplyr::select(Gene.name, sid, bh) %>% group_by(Gene.name) %>% top_n(-1, bh) %>% ungroup() %>% dplyr::select(Gene.name, sid)

option_list = list(
    make_option(c("-M", "--molNom"), action="store", default=NA, type='character', help="molecular Nom results"),
    make_option(c("-O", "--output"), action="store", default=NA, type='character', help="output file for total APA sig snps in mol qtl")
)

opt_parser <- OptionParser(option_list=option_list)
opt <- parse_args(opt_parser)


if (opt$molNom == "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_prot.fixed.nominal.out") {
  in_file=read.table(opt$molNom, col.names = c("Gene.stable.ID", "sid", "dist", "pval", "slope"),stringsAsFactors=F)
  file_newNames=in_file %>%  inner_join(geneNames, by="Gene.stable.ID") %>% dplyr::select("Gene.name", "sid", "pval")
} else {
in_file=read.table(opt$molNom, col.names = c("pid", "sid", "dist", "pval", "slope"),stringsAsFactors=F)
file_newNames=in_file %>% separate(pid, into=c("Gene.stable.ID", "ver"), sep ="[.]") %>% inner_join(geneNames, by="Gene.stable.ID") %>% dplyr::select("Gene.name", "sid", "pval")
}

overlap= file_newNames %>% semi_join(sigSNPgene, by=c("Gene.name", "sid")) 

write.table(overlap, file=opt$output, quote=F, col.names = T, row.names = F)



``` 

Run this first on the rnaQTLs.  


run_sigTotAPAinMolPhenoRNA.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPhenoRNA
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPhenoRNA.out
#SBATCH --error=run_sigTotAPAinMolPhenoRNA.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_RNAseq_phase2.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPvalRNA.txt" 
```




run_sigTotAPAinMolPhenoProt.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPhenoProt
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPhenoProt.out
#SBATCH --error=run_sigTotAPAinMolPhenoProt.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_prot.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPvalProtein.txt" 

```



run_sigTotAPAinMolPhenoProt.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPhenoProt
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPhenoProt.out
#SBATCH --error=run_sigTotAPAinMolPhenoProt.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_prot.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPvalProtein.txt" 

```



run_sigTotAPAinMolPhenoRNAg.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPhenoRNAg
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPhenoRNAg.out
#SBATCH --error=run_sigTotAPAinMolPhenoRNAg.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_RNAseqGeuvadis.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPvalRNAg.txt" 

```

run_sigTotAPAinMolPhenoRibo.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPhenoRibo
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPhenoRibo.out
#SBATCH --error=run_sigTotAPAinMolPhenoRibo.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_ribo_phase2.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPvalribo.txt" 

```

run_sigTotAPAinMolPheno4su.sh

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigTotAPAinMolPheno4su
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigTotAPAinMolPheno4su.out
#SBATCH --error=run_sigTotAPAinMolPheno4su.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_4su30.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPval4su30.txt" 

Rscript sigTotAPAinMolPheno.R --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_4su60.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molTotal/TotAPAqtlsPval4su60.txt" 
```



##MolQTL pvalues for Nuclear ApaQTLs    


sigNucAPAinMolPheno.R  
```{r,eval=F}
#!/bin/rscripts

#this script creates takes in the permuted APAQTL results for the total fraction and nominal pvalues from the molecular phenotpye  molecular phenotype 

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(optparse)

geneNames=read.table("/project2/gilad/briana/genome_anotation_data/ensemble_to_genename.txt", sep="\t", header=T, stringsAsFactors = F)

nuc_perm=read.table("/project2/gilad/briana/threeprimeseq/data/perm_APAqtl_trans/filtered_APApeaks_merged_allchrom_refseqGenes_pheno_Nuclear_transcript_permResBH.txt", header = T,stringsAsFactors=F)

sigSNPgene=nuc_perm %>% filter(-log10(bh)>1) %>% separate(pid, into=c("chr", "start", "end", "id"), sep=":") %>% separate(id, into=c("Gene.name", "strand", "peaknum"), sep="_") %>% dplyr::select(Gene.name, sid, bh) %>% group_by(Gene.name) %>% top_n(-1, bh) %>% ungroup() %>% dplyr::select(Gene.name, sid)

option_list = list(
    make_option(c("-M", "--molNom"), action="store", default=NA, type='character', help="molecular Nom results"),
    make_option(c("-O", "--output"), action="store", default=NA, type='character', help="output file for total APA sig snps in mol qtl")
)

opt_parser <- OptionParser(option_list=option_list)
opt <- parse_args(opt_parser)


if (opt$molNom == "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_prot.fixed.nominal.out") {
  in_file=read.table(opt$molNom, col.names = c("Gene.stable.ID", "sid", "dist", "pval", "slope"),stringsAsFactors=F)
  file_newNames=in_file %>%  inner_join(geneNames, by="Gene.stable.ID") %>% dplyr::select("Gene.name", "sid", "pval")
} else {
in_file=read.table(opt$molNom, col.names = c("pid", "sid", "dist", "pval", "slope"),stringsAsFactors=F)
file_newNames=in_file %>% separate(pid, into=c("Gene.stable.ID", "ver"), sep ="[.]") %>% inner_join(geneNames, by="Gene.stable.ID") %>% dplyr::select("Gene.name", "sid", "pval")
}

overlap= file_newNames %>% semi_join(sigSNPgene, by=c("Gene.name", "sid")) 

write.table(overlap, file=opt$output, quote=F, col.names = T, row.names = F)

``` 

1 bash script for all of the phenotypes

run_sigNucAPAinMolPheno.sh  

```{bash,eval=F}
#!/bin/bash


#SBATCH --job-name=run_sigNucAPAinMolPheno
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=run_sigNucAPAinMolPheno.out
#SBATCH --error=run_sigNucAPAinMolPheno.err
#SBATCH --partition=bigmem2
#SBATCH --mem=64G
#SBATCH --mail-type=END

module load R 


Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_RNAseq_phase2.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPvalRNA.txt" 


Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_prot.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPvalProtein.txt"  

Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_RNAseqGeuvadis.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPvalRNAg.txt" 


Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_ribo_phase2.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPvalribo.txt" 



Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_4su30.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPval4su30.txt" 

Rscript run_sigNucAPAinMolPheno.sh --molNom "/project2/gilad/briana/threeprimeseq/data/molecular_QTLs/nom/fastqtl_qqnorm_4su60.fixed.nominal.out" --output "/project2/gilad/briana/threeprimeseq/data/molecular_overlap/APA2molNuclear/NucAPAqtlsPval4su60.txt" 


```






##Create Histograms  

I will next estimate sharing with pi_1 and create histograms of the resulting pvalues.  
* Protein  

```{r}
totAPAinProt=read.table("../data/mol_overlap/APA2molTotal/TotAPAqtlsPvalProtein.txt", header = T, stringsAsFactors = F)
qval_prot=pi0est(totAPAinProt$pval, pi0.method = "bootstrap")
hist(totAPAinProt$pval)
text(.8,10, paste("pi_1=", round((1-qval_prot$pi0), digit=3), sep=" "))

```
* RNA  

```{r}
totAPAinRNA=read.table("../data/mol_overlap/APA2molTotal/TotAPAqtlsPvalRNA.txt", header = T, stringsAsFactors = F)
qval_RNA=pi0est(totAPAinRNA$pval, pi0.method = "bootstrap")
hist(totAPAinRNA$pval)
text(.8,18, paste("pi_1=", round((1-qval_RNA$pi0), digit=3), sep=" "))

```

