---
title: "Peak QC data and figures"
author: "Briana Mittleman"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to do some QC and filtering on the peaks to go along with the number of peaks to cover % of a gene figure.  


* Number of called peaks  

* peaks used at X% in total/nuclear  

* number of genes  


```{r}
library(tidyverse)
library(data.table)
library(workflowr)
library(cowplot)
```

```{r}
totalPeakUs=read.table("../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total.pheno_fixed.txt.gz", header = T, stringsAsFactors = F) %>% separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% separate(id, sep="_", into=c("gene", "strand", "peak"))
nuclearPeakUs=read.table("../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear.pheno_fixed.txt.gz", header = T, stringsAsFactors = F) %>% separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% separate(id, sep="_", into=c("gene", "strand", "peak"))
```

There are 338141 called peaks in the data. 

I need to make the fractions numeric, I will do this in python because I can go through each value, split them and get the numeric.  


It will be easiest if I write the counts out:  

```{r}
#write.table(totalPeakUs[,7:45], file="../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total.pheno_fixed.CountsOnly",quote=FALSE, col.names = F, row.names = F)

#write.table(nuclearPeakUs[,7:45], file="../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear.pheno_fixed.CountsOnly",quote=FALSE, col.names = F, row.names = F)

```

Move these to /project2/gilad/briana/threeprimeseq/data/PeakUsage

convertCount2Numeric.py
```{bash,eval=F}
def convert(infile, outfile):
  final=open(outfile, "w")
  for ln in open(infile, "r"):
    line_list=ln.split()
    new_list=[]
    for i in line_list:
      num, dem = i.split("/")
      if dem == "0":
        perc = "0.00"
      else:
        perc = int(num)/int(dem)
        perc=round(perc,2)
        perc= str(perc)
      new_list.append(perc)
    final.write("\t".join(new_list)+ '\n')
  final.close()
  
convert("/project2/gilad/briana/threeprimeseq/data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total.pheno_fixed.CountsOnly","/project2/gilad/briana/threeprimeseq/data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total.pheno_fixed.CountsOnlyNUMERIC.txt" )


convert("/project2/gilad/briana/threeprimeseq/data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear.pheno_fixed.CountsOnly","/project2/gilad/briana/threeprimeseq/data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear.pheno_fixed.CountsOnlyNUMERIC.txt")

```


Because any value less than .001 becomes 0, all peaks for a gene will not add to zero.  


```{r}
ind=colnames(totalPeakUs)[7:dim(totalPeakUs)[2]]
totalPeakUs_CountNum=read.table("../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total.pheno_fixed.CountsOnlyNUMERIC.txt", col.names = ind)

nuclearPeakUs_CountNum=read.table("../data/PeakUsage/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear.pheno_fixed.CountsOnlyNUMERIC.txt", col.names = ind)
```

Numeric values with the annotations:  

```{r}
totalPeak=as.data.frame(cbind(totalPeakUs[,1:6], totalPeakUs_CountNum))
nuclearPeak=as.data.frame(cbind(nuclearPeakUs[,1:6], nuclearPeakUs_CountNum))
```

Get the mean coverage for each peak.

```{r}
totalPeakUs_CountNum_mean=rowMeans(totalPeakUs_CountNum)
nuclearPeakUs_CountNum_mean=rowMeans(nuclearPeakUs_CountNum)
```

Append these to the inforamtion about the peak.   


```{r}
TotalPeakUSMean=as.data.frame(cbind(totalPeakUs[,1:6],totalPeakUs_CountNum_mean))
NuclearPeakUSMean=as.data.frame(cbind(nuclearPeakUs[,1:6],nuclearPeakUs_CountNum_mean))
                              
```


Get the number of genes with mean(usage > 5%)  

Total:  
```{r}
TotalPeakUSMean_filt=TotalPeakUSMean %>% filter(totalPeakUs_CountNum_mean>=.05) %>% group_by(gene) %>% summarise(Npeaks=n())

```

I want to get how many genes have 1,2,3,4 ect:  

```{r}
totalPeaksPerGene=TotalPeakUSMean_filt %>% group_by(Npeaks) %>% summarise(GenesWithNPeaks=n())

ggplot(totalPeaksPerGene,aes(x=Npeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity",fill="darkviolet") + labs(x="Number Peaks with >5% usage", y="Number of Genes", title="Genes with peaks covering > 5% in Total")
```



Nuclear:  
```{r}
NuclearPeakUSMean_filt=NuclearPeakUSMean %>% filter(nuclearPeakUs_CountNum_mean>=.05) %>% group_by(gene) %>% summarise(Npeaks=n())

```

I want to get how many genes have 1,2,3,4 ect:  

```{r}
nuclearPeaksPerGene=NuclearPeakUSMean_filt %>% group_by(Npeaks) %>% summarise(GenesWithNPeaks=n())
nuclearPeaksPerGene$GenesWithNPeaks=as.integer(nuclearPeaksPerGene$GenesWithNPeaks)
ggplot(nuclearPeaksPerGene,aes(x=Npeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity", fill="deepskyblue3") + labs(x="Number Peaks with >5% usage", y="Number of Genes", title="Genes with peaks covering > 5% in Nuclear")
```

Genes with at least 1: 

```{r}
#nuclear
nrow(NuclearPeakUSMean_filt)  

#total
nrow(TotalPeakUSMean_filt)  

```

Join them to put on the same plot:  
gene level

```{r}
nPeaksBoth_gene=TotalPeakUSMean_filt %>% full_join(NuclearPeakUSMean_filt, by="gene")
colnames(nPeaksBoth_gene)= c("Gene", "Total", "Nuclear")
nPeaksBoth_gene$Nuclear= nPeaksBoth_gene$Nuclear %>% replace_na(0)
nPeaksBoth_gene$Total= nPeaksBoth_gene$Total %>% replace_na(0)
nPeaksBoth_gene=nPeaksBoth_gene %>% mutate(Difference=Nuclear-Total)

ggplot(nPeaksBoth_gene, aes(x=Difference)) + geom_histogram() + labs(title="Distribution of difference in number of \n Peaks >5% between Nuclear and Total", y="Genes", x="Nuclear - Total")
summary(nPeaksBoth_gene$Difference)
```


WHich genes is this most affecting:
```{r}
nPeaksBoth_gene %>% arrange(desc(Difference)) %>% slice(1:10)
```





**Look at some of these in IGV**  

Good examples: API5,ARL3, SHQ1




peak number level: 
```{r}
nPeaksBoth=totalPeaksPerGene %>% full_join(nuclearPeaksPerGene, by="Npeaks")
colnames(nPeaksBoth)= c("Peaks", "Total", "Nuclear")
nPeaksBoth$Total= nPeaksBoth$Total %>% replace_na(0)

#melt nPeaksBoth
nPeaksBoth_melt=melt(nPeaksBoth, id.var="Peaks")
colnames(nPeaksBoth_melt)= c("Peaks", "Fraction", "Genes")
```

Make a plot:

```{r}
peakUsage5perc=ggplot(nPeaksBoth_melt, aes(x=Peaks, y=Genes, fill=Fraction)) + geom_bar(stat="identity", position = "dodge") + labs(title="Number of Genes with >5% Peak Usage") + theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  + facet_grid(~Fraction)

peakUsage5perc

ggsave(peakUsage5perc, file="../output/plots/QC_plots/peakUsage5perc.png")
```


Peaks with >5 per not at gene level:  


```{r}
#nuclear  
NuclearPeakUSMean %>% filter(nuclearPeakUs_CountNum_mean>=.05) %>% nrow()

#total
TotalPeakUSMean %>% filter(totalPeakUs_CountNum_mean>=.05) %>% nrow()
```



Plot distributions priots to filtering:  
```{r}
NuclearPeakUSMean_sm=NuclearPeakUSMean %>% select(peak, nuclearPeakUs_CountNum_mean)
TotalPeakUSMean_sm=TotalPeakUSMean %>% select(peak, totalPeakUs_CountNum_mean)
BothPeakUSMean=TotalPeakUSMean_sm %>% full_join(NuclearPeakUSMean_sm, by=c("peak"))
summary(BothPeakUSMean)

```


```{r}
colnames(BothPeakUSMean)=c("Peak", "Total", "Nuclear")
BothPeakUSMean_melt=melt(BothPeakUSMean, id.vars = "Peak")
colnames(BothPeakUSMean_melt)=c("Peak", "Fraction", "MeanUsage")
meanUsBox=ggplot(BothPeakUSMean_melt,aes(y=MeanUsage, x=Fraction, fill=Fraction)) +geom_boxplot() +scale_fill_manual(values=c("darkviolet","deepskyblue3"))
meanUsBoxZoom=ggplot(BothPeakUSMean_melt,aes(y=MeanUsage, x=Fraction, fill=Fraction)) +geom_boxplot() +ylim(c(0,.05))+scale_fill_manual(values=c("darkviolet","deepskyblue3"))

meanUsBoxBoth=plot_grid(meanUsBox,meanUsBoxZoom)

ggsave(file="../output/plots/QC_plots/meanPeakUsageBoxPlots.png",meanUsBoxBoth)
```
```{r}

meanUs_den=ggplot(BothPeakUSMean_melt,aes(x=MeanUsage, by=Fraction, fill=Fraction)) +geom_density(alpha=.4) +scale_fill_manual(values=c("darkviolet","deepskyblue3"))
meanUs_denZoom=ggplot(BothPeakUSMean_melt,aes(x=MeanUsage, by=Fraction, fill=Fraction)) +geom_density(alpha=.4) +xlim(c(0,.05)) + scale_fill_manual(values=c("darkviolet","deepskyblue3"))

meanUs_denBoth=plot_grid(meanUs_den,meanUs_denZoom)

ggsave(file="../output/plots/QC_plots/meanPeakUsagDensityPlots.png",meanUs_denBoth,)
```





With means at about 4%. I may remake these plots with 1 %  

```{r}
TotalPeakUSMean_filt1=TotalPeakUSMean %>% filter(totalPeakUs_CountNum_mean>=.01) %>% group_by(gene) %>% summarise(Npeaks=n())

totalPeaksPerGene1=TotalPeakUSMean_filt1 %>% group_by(Npeaks) %>% summarise(GenesWithNPeaks=n())

ggplot(totalPeaksPerGene1,aes(x=Npeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity",fill="darkviolet") + labs(x="Number Peaks with >1% usage", y="Number of Genes", title="Genes with peaks covering > 1% in Total")

```

Nuclear:  
```{r}
NuclearPeakUSMean_filt1=NuclearPeakUSMean %>% filter(nuclearPeakUs_CountNum_mean>=.01) %>% group_by(gene) %>% summarise(Npeaks=n())

```

I want to get how many genes have 1,2,3,4 ect:  

```{r}
nuclearPeaksPerGene1=NuclearPeakUSMean_filt1 %>% group_by(Npeaks) %>% summarise(GenesWithNPeaks=n())
nuclearPeaksPerGene1$GenesWithNPeaks=as.integer(nuclearPeaksPerGene1$GenesWithNPeaks)
ggplot(nuclearPeaksPerGene1,aes(x=Npeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity", fill="deepskyblue3") + labs(x="Number Peaks with >1% usage", y="Number of Genes", title="Genes with peaks covering > 1% in Nuclear")
```


Try to do this with 5% in 2/3 of the libraries instead:  

```{r}
keep.exprs_T=rowSums(as.matrix(totalPeakUs_CountNum>=.05)) >= 26
TotalPeakUS_filt= as.data.frame(cbind(totalPeakUs[,1:6], totalPeakUs_CountNum))[keep.exprs_T,]



keep.exprs_N=rowSums(as.matrix(nuclearPeakUs_CountNum>=.05)) >= 26
nuclearPeakUS_filt= as.data.frame(cbind(nuclearPeakUs[,1:6],nuclearPeakUs_CountNum))[keep.exprs_N,]


```
Total: 30657 peaks pass this filter
Nuclear:44030 peaks pass 



Now I can group by gene and see what happens:  

```{r}
TotalPeakUS_filt_gene=TotalPeakUS_filt %>% group_by(gene) %>% summarise(nPeaks=n())

TotalPeaksPerGene2=TotalPeakUS_filt_gene %>% group_by(nPeaks) %>% summarise(GenesWithNPeaks=n())
ggplot(TotalPeaksPerGene2,aes(x=nPeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity", fill="darkviolet") + labs(x="Number Peaks with >5% usage in 2/3 ind", y="Number of Genes", title="Genes with peaks covering > 5% in 2/3 Ind Total")

```

In this filter there are 13497 genes. 



```{r}
nuclearPeakUS_filt_gene=nuclearPeakUS_filt %>% group_by(gene) %>% summarise(nPeaks=n())

nuclearPeaksPerGene2=nuclearPeakUS_filt_gene %>% group_by(nPeaks) %>% summarise(GenesWithNPeaks=n())
ggplot(nuclearPeaksPerGene2,aes(x=nPeaks,y=GenesWithNPeaks)) + geom_bar(stat="identity", fill="deepskyblue3") + labs(x="Number Peaks with >5% usage in 2/3 ind", y="Number of Genes", title="Genes with peaks covering > 5% in 2/3 Ind Nuclear")
```
14690

Melt these to put on one plot:  
```{r}
nPeaksBoth_filt2=TotalPeaksPerGene2 %>% full_join(nuclearPeaksPerGene2, by="nPeaks")
colnames(nPeaksBoth_filt2)= c("Peaks", "Total", "Nuclear")
nPeaksBoth_filt2$Total= nPeaksBoth_filt2$Total %>% replace_na(0)

#melt nPeaksBoth
nPeaksBoth_melt2=melt(nPeaksBoth_filt2, id.var="Peaks")
colnames(nPeaksBoth_melt2)= c("Peaks", "Fraction", "Genes")
```

Plot:  

```{r}
peakUsage5perc2=ggplot(nPeaksBoth_melt2, aes(x=Peaks, y=Genes, fill=Fraction)) + geom_bar(stat="identity", position = "dodge") + labs(title="Number of Genes with >5% peak usage in 2/3 Ind") + theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("darkviolet","deepskyblue3"))  + facet_grid(~Fraction)

peakUsage5perc2

ggsave(peakUsage5perc2, file="../output/plots/QC_plots/peakUsage5perc2.3ind.png")
```


```{r}
nPeaksBoth2_gene=TotalPeakUS_filt_gene %>% full_join(nuclearPeakUS_filt_gene, by="gene")


colnames(nPeaksBoth2_gene)= c("Gene", "Total", "Nuclear")
nPeaksBoth2_gene$Nuclear= nPeaksBoth2_gene$Nuclear %>% replace_na(0)
nPeaksBoth2_gene$Total= nPeaksBoth2_gene$Total %>% replace_na(0)
nPeaksBoth2_gene=nPeaksBoth2_gene %>% mutate(Difference=Nuclear-Total)

PeakDiffPlot_5perc2.3ind=ggplot(nPeaksBoth2_gene, aes(x=Difference)) + geom_histogram() + labs(title="Distribution of difference in number of \n Peaks >5% Usage in 2/3 Individuals \n between Nuclear and Total", y="Genes", x="Nuclear - Total")
summary(nPeaksBoth2_gene$Difference)

ggsave(file="../output/plots/QC_plots/PeakDiffPlot_5perc2.3ind.png", PeakDiffPlot_5perc2.3ind)
```

The mean is this set is .9
