---
title: "Exploritory analysis prior to running ash"
author: "Briana Mittleman"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are a few things about the data I need to understand before I can run ash. First I need to find the genes that overlap with protein and RNA. Then I need to pick those with 1 dominant peak. 

```{r}
set.seed(1)
library(workflowr)
library(tidyverse)
library(reshape2)
library(cowplot)
```


Upload data: 




I want the filtered peak counts. I need to filter the counts file for the total fraction based on the filtered peaks. 

Total counts:  /project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov_noMP/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_fixed.fc

okPeaks: /project2/gilad/briana/threeprimeseq/data/PeakUsage_noMP/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_fixed.pheno.5percPeaks.txt  


filterTotalCounts_noMP_5percCov.py
```{bash,eval=F}
totalokPeaks5perc_file="/project2/gilad/briana/threeprimeseq/data/PeakUsage_noMP/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_fixed.pheno.5percPeaks.txt"
countFile=open("/project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov_noMP/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_fixed.fc","r")

outFile=open("/project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov_noMP/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_5percCov_fixed.fc", "w")

allPeakOk={}
for ln in open(totalokPeaks5perc_file,"r"):
    peakname=ln.split()[5]
    peaknum=peakname[4:]
    if peaknum not in allPeakOk.keys():
        allPeakOk[peaknum]=""
    
for i,ln in enumerate(countFile):
    if i==1:
      outFile.write(ln)
    if i>1:
      ID=ln.split()[0]
      peak=ID.split(":")[0]
      peaknum=peak[4:]
      if peaknum in allPeakOk.keys():
          outFile.write(ln)
outFile.close()
```



```{r}
total_Cov=read.table("../data/PeakCounts_noMP_5perc/filtered_APApeaks_merged_allchrom_refseqGenes.TranscriptNoMP_sm_quant.Total_5percCov_fixed.fc", stringsAsFactors = F,header = T) %>%  separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "Name"), sep=":")
total_genes=total_Cov %>% select(Name) %>% arrange(Name) %>% unique()
```
Gene names:  

```{r}
geneNames=read.table("../data/ensemble_to_genename.txt",sep="\t", header=T, stringsAsFactors = F, col.names=c("ID", "Name", "Source"))
```

```{r}
prot=read.table("../data/mol_pheno/fastqtl_qqnorm_prot.fixed.noChr.txt",header=T,stringsAsFactors = F) %>% inner_join(geneNames, by="ID")
```

Keep the protein genes in APA:  

```{r}
prot_inAPA=prot %>% semi_join(total_genes, by="Name")
```


This shows we have 4209 genes with data for both. Now I can back filter the total peaks for the genes in prot_inAPA  

```{r}
total_Cov_wProt= total_Cov %>% semi_join(prot_inAPA,by="Name")
```

Need to give Stephens lab:  unadjusted R-squared, and n and p for each protein, where p is the number of apa s that you are using in the regression and n is the number of samples?

To do this I need to get the overlapping individuals:  

```{r}
protInd=colnames(prot)[5:(dim(prot)[2]-2)]

ApaInd=c()
for (i in colnames(total_Cov)[12:ncol(total_Cov)]){
  num=substr(i,2,6)
  name=paste("NA", num, sep="")
  ApaInd=c(ApaInd, name)
}


IndBoth=intersect(protInd,ApaInd)
length(IndBoth)
```

I have 29 individuals in common for these. 


I need to make a matrix for each gene. It will have a row for each commmon individual. A column for the protein, and a column for each assocaited peaks. After I have this I will be able to get the R2 value.  


First create a function.

```{r}
get_R2=function(gene, Cov=total_Cov, prot=prot_inAPA, apaName=ApaInd){
  gene_un= enexpr(gene)
  #deal with APA
  genePeaks=total_Cov %>% filter(Name==!!gene_un)
  n=nrow(genePeaks)
  drop_col=c('chr','Chr', 'start','end','strand','Name', 'Start','End','Strand','Length')
  genePeaks_sm= genePeaks %>% select(-one_of(drop_col))
  colnames(genePeaks_sm)=c("peak", ApaInd) 
  genePeakM=genePeaks_sm %>% column_to_rownames(var="peak") %>% t()
  genePeakDF=as.data.frame(genePeakM) %>% rownames_to_column(var="Ind")
  #deal with prot
  drop_col_prot= c("Chr", "start", "end", "ID", "Name", "Source")
  geneProt=prot %>% filter(Name==!!gene_un) %>% select(-one_of(drop_col_prot)) %>% t()
  colnames(geneProt)="prot"
  #print(dim(geneProt))
  geneProt_df=as.data.frame(geneProt) %>% rownames_to_column(var="Ind") %>% drop_na(prot)
  #print(geneProt_df)
  both=geneProt_df %>% inner_join(genePeakDF,by="Ind")
  num=seq(1,n)
  base="summary(lm(both$prot~"
  for (i in 3:dim(both)[2]){
    base=paste(base, "+both[,",i,"]",sep="")
  }
  code=paste(base, "))$r.squared", sep="")
  r2=eval(parse(text=code))
  final=c(gene, r2,nrow(both),n)
  return(final)
}

```

Run this on all genes:

```{r}
final_matrix=matrix(c("gene","r2","n","p"),1,4)
for (i in prot_inAPA$Name){
 final_matrix= rbind(final_matrix,get_R2(i))
}


```

Make this a dataframe:  
```{r}
final_df=as.data.frame(final_matrix)

colnames(final_df)=as.character(unlist(final_df[1,]))

final_df <- final_df[-1 ,]

save(final_df,file="../data/protAndAPAlmRes.Rda")
```

