---
title: "Distal vs Proximal Peak DeepTools"
author: "Briana Mittleman"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to show RNA seq vs 3' seq in peaks that are internal. 

To do this I need to pull in the peaks (use the filtered ones from peakQCPlots) to get peaks that are used but are not the most distal for the gene. I need to split by strand when i do this becasue most distal is different in both cases. 
 
 
 
Load Libraries
```{r}
library(data.table)
library(tidyverse)
library(workflowr)
library(cowplot)
```



Load peaks for total and nuclear post filter:  

```{r}
total_PeakUsMean=read.table("../data/PeakUsage/PeakUsageMeanPost5percFilter.Total.txt", head = T)
nuclear_PeakUsMean=read.table("../data/PeakUsage/PeakUsageMeanPost5percFilter.Nuclear.txt", head=T)
```

Seperate positive and negative:

* Total  

```{r}
total_PeakUsMean_pos=total_PeakUsMean %>% filter(strand=="+")
total_PeakUsMean_neg=total_PeakUsMean %>% filter(strand=="-")
```



* Nuclear  

```{r}
nuclear_PeakUsMean_pos=nuclear_PeakUsMean %>% filter(strand=="+")
nuclear_PeakUsMean_neg=nuclear_PeakUsMean %>% filter(strand=="-")
```


Group by gene and keep internal (remove genes with only 1)

* For positive strand keep bottom one with top_n() peak #  (most internal)

```{r}
total_PeakUsMean_pos_internal=total_PeakUsMean_pos %>% group_by(gene) %>% mutate(n=n()) %>% filter(n>1) %>% top_n(1,peak)

nuclear_PeakUsMean_pos_internal=nuclear_PeakUsMean_pos %>% group_by(gene) %>% mutate(n=n()) %>% filter(n>1) %>% top_n(1,peak)
```


*For negative strand use top_n(-1)   


```{r}
total_PeakUsMean_neg_internal=total_PeakUsMean_neg %>% group_by(gene) %>% mutate(n=n()) %>% filter(n>1) %>% top_n(-1,peak)

nuclear_PeakUsMean_neg_internal=nuclear_PeakUsMean_neg %>% group_by(gene) %>% mutate(n=n()) %>% filter(n>1) %>% top_n(-1,peak)
```


Bind the total and nuclear rows back together:  

Total  

```{r}
total_PeakUsMean_internal=as.data.frame(rbind(total_PeakUsMean_pos_internal,total_PeakUsMean_neg_internal)) %>% arrange(chr, start, end)
```

Nuclear:  
```{r}
nuclear_PeakUsMean_internal=as.data.frame(rbind(nuclear_PeakUsMean_pos_internal,nuclear_PeakUsMean_neg_internal)) %>% arrange(chr, start, end)
```


Use these peak numbers to filter the bed file that I use for deep tools. I can do this in python with a dictionary of the peaks to keep. 
