---
title: "Location of Signal Site"
author: "Briana Mittleman"
date: "3/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the [Signal Site enrichment analysis](SignalSiteEnrich.html) I looked at the peaks to see if signal sites are enriched upstream of my peaks. I found this is true but now I want to see where the signal sites are in comparison to my peaks. I am going to use the biostrings package tool matchPWM for this analysis.
```{r}
library(workflowr)
library(tidyverse)
library(Biostrings)
library(BSgenome)
library(genomation)

```
##Make location file
I need to get the coordinates for the regions I care about. I want to look at the peak and 150bp upstream. This is probably larger than I will need  to look at but it will be good to have an inclusive look first.  


I want to use the peak file and make a file that is the peak and upstream 150:  


Upstream150Bases.py
```{python, eval=F}
#python  
def main(Fin, Fout):
  outBed=open(Fout, "w")
  chrom_lengths=open("/project2/gilad/briana/genome_anotation_data/chrom_lengths2.sort.bed","r")
  #make a dictionary with chrom lengths
  length_dic={}
  for i in chrom_lengths:
    chrom, start, end = i.split()
    length_dic[str(chrom)]=int(end)  

#write file 
  for ln in open(Fin):
    chrom, start, end, name, score, strand = ln.split()
    chrom=str(chrom)
    if strand=="+":
      start_new=int(start)-150
      if start_new <= 1:
        start_new = 0 
      end_new= int(end)
      if end_new == 0:
        end_new=1
      outBed.write("%s\t%d\t%d\t%s\t%s\t%s\n"%(chrom, start_new, end_new, name, score, strand))
    if strand == "-":
      start_new=int(start)
      end_new=int(end) + 150
      outBed.write("%s\t%d\t%d\t%s\t%s\t%s\n"%(chrom, start_new, end_new, name, score, strand))
  outBed.close()  

if __name__ == "__main__":
    import sys
    inFile = sys.argv[1]
    outFile=sys.argv[2] 
    main(inFile, outFile)
```



run_get150up.sh
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=run_get150up
#SBATCH --account=pi-yangili1
#SBATCH --time=36:00:00
#SBATCH --output=run_get150upt.out
#SBATCH --error=run_get150up.err
#SBATCH --partition=broadwl
#SBATCH --mem=16G
#SBATCH --mail-type=END

module load Anaconda3
source activate three-prime-env

python Upstream150Bases.py /project2/gilad/briana/threeprimeseq/data/peaks4DT/APAPeaks_5percCov_fixedStrand.bed  /project2/gilad/briana/threeprimeseq/data/Signal_Loc/APAPeaks_5percCov_fixedStrand_peakand150up.bed 
```

##Get subject (reads)

Input the regions: 


Fix chromosomes: 
```{r}
PeakRegions=read.table("../data/Signal_Loc/APAPeaks_5percCov_fixedStrand_peakand150up.bed", header=F,col.names = c("chr","start", "end", "peak", "score", "strand")) %>% mutate(Chrom=paste("chr", chr, sep="")) %>% select(Chrom, start,end,peak,score,strand)
write.table(PeakRegions, file="../data/Signal_Loc/APAPeaks_5percCov_fixedStrand_peakand150up_fixedChr.bed", quote=F, col.names = F, row.names = F, sep="\t")


```



```{r}
#convert to reads 

reads.GR= readGeneric(file="../data/Signal_Loc/APAPeaks_5percCov_fixedStrand_peakand150up_fixedChr.bed",chr =1, start = 2, end =3,  meta.cols =4, header=F, zero.based=TRUE,strand=6)


```

I need to overlap these positions with the genome

###Make motifs  

```{r}
AATAAA= PWM("AATAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
ATTAAA= PWM("ATTAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
AGTAAA= PWM("AGTAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
TATAAA= PWM("TATAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
CATAAA= PWM("CATAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
GATAAA= PWM("GATAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
AATATA= PWM("AATATA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
AATACA= PWM("AATACA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
AATAGA= PWM("AATAGA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
AAAAAG= PWM("AAAAAG", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
ACTAAA= PWM("ACTAAA", type = c("log2probratio", "prob"), prior.params = c(A=0.25, C=0.25, G=0.25, T=0.25))
```



###find the mathes  

```{r}
genome.hg19 <- getBSgenome("BSgenome.Hsapiens.UCSC.hg19")

#matches <- matchPWM(pwm=AATAAA, subject = genome.hg19) %>% keepStandardChromosomes(., species= "Homo sapiens")


DNAstringSetPeaks=data.frame(seq=getSeq(genome.hg19, reads.GR))

x=DNAString(DNAstringSetPeaks[1,1])

hits <- matchPWM(AATAAA,x,with.score=T) 


start(hits)
```



Look over and make hits file for all 

```{r}
list_AATAAA_res=c()
for (i in 1:nrow(DNAstringSetPeaks)){
  x=DNAString(DNAstringSetPeaks[i,1])
  list_AATAAA_res=c(list_AATAAA_res,matchPWM(AATAAA,x,with.score=T))
}
```


Get out the start positions: 


```{r}
starts_AATAAA=c()
nsig=c()
last_oc_AATAAA=c()
for (i in list_AATAAA_res){
  nsig=c(nsig, length(start(i)))
  starts_AATAAA=c(starts_AATAAA, start(i))
  #print(length(start(i)))
  if (length(start(i)) != 0 ){
    last_oc_AATAAA=c(last_oc_AATAAA, max(start(i),na.rm =T))
  }
}
   
```




Histogram of results:

```{r}
summary(starts_AATAAA)
hist(starts_AATAAA,breaks=10000)
```

```{r}
summary(nsig)
sum(nsig==0)
sum(nsig==1)

hist(nsig,breaks=100)
```

Look at the first occurence:  

```{r}
summary(last_oc_AATAAA)
hist(last_oc_AATAAA,breaks=1000)
```

I want to get the closest occurance of the 
