---
title: "Understand Peaks"
author: "Briana Mittleman"
date: "12/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this analysis is to understand the data a bit better at the peak level. I want to have the cleanest set of peaks when I perform the final anlyses for the paper.

##Variation in peaks  

First I will run PCA on the peak coverage. I will run this seperatly for the total and nuclear fractions. I do not expect large amount of separation.  


I will use the peak coverage data before the ratios are created for leafcutter. These files were created using feature counts on the filtered peaks. At this point the peaks have been mapped to the closest refseq transcript on the opposite strand. 

Relevant file:  
* /project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc  

* /project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc  

These files are in /Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/PeakCounts on my computer.  

```{r}
library(tidyverse)
library(workflowr)
library(cowplot)
library(reshape2)
library(devtools)
library(tximport)

```

Load data:  

```{r}
#only keep the counts 
total_Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,7:45]
nuclear_Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,7:45]

```

```{r}
ggplot(total_Cov, aes(x=log10(X18486_T))) + geom_density()
```


###Total:  
Run PCA on the total coverage 

```{r}
pca_tot_peak=prcomp(total_Cov, center=T,scale=T)
summary(pca_tot_peak)
pca_tot_df=as.data.frame(pca_tot_peak$rotation) %>% rownames_to_column(var="lib") %>% mutate(line=substr(lib,2,6))
pca_tot_df$line=as.integer(pca_tot_df$line)
```

I want to color these by library size. 

```{r}
map_stats=read.csv("../data/comb_map_stats_39ind.csv", header=T)

map_stat_total=map_stats %>% filter(fraction=="total")
map_stat_total$batch=as.factor(map_stat_total$batch)
```

Join the relevant stats with the pca dataframe.  

```{r}
pca_tot_df=pca_tot_df %>% full_join(map_stat_total, by="line")
```

Plot this PCA:  

```{r}
totPCA_batch=ggplot(pca_tot_df, aes(x=PC1, y=PC2, col=batch )) + geom_point() + labs(x="PC1:0.89", y="PC2:0.043", title="Raw PAS qunatification data Total \n colored by batch ")
ggsave("../output/plots/QC_plots/TotalPCA_colBatch.png",totPCA_batch)

totPCA_mapped=ggplot(pca_tot_df, aes(x=PC1, y=PC2, col=comb_mapped )) + geom_point() + labs(x="PC1:0.89", y="PC2:0.043", title="Raw PAS qunatification data Total \n colored by Mapped Read count")
ggsave("../output/plots/QC_plots/TotalPCA_colMapped.png",totPCA_mapped)
```


###Nuclear  

Run PCA on the Nuclear coverage 

```{r}
pca_nuc_peak=prcomp(nuclear_Cov, center=T,scale=T)
summary(pca_nuc_peak)
pca_nuc_df=as.data.frame(pca_nuc_peak$rotation) %>% rownames_to_column(var="lib") %>% mutate(line=substr(lib,2,6))
pca_nuc_df$line=as.integer(pca_nuc_df$line)
```

I want to color these by library size. 

```{r}
map_stat_nuclear=map_stats %>% filter(fraction=="nuclear")
map_stat_nuclear$batch=as.factor(map_stat_nuclear$batch)
```

Join the relevant stats with the pca dataframe.  

```{r}
pca_nuc_df=pca_nuc_df %>% full_join(map_stat_nuclear, by="line")
```

Plot this PCA:  

```{r}
nucPCA_batch=ggplot(pca_nuc_df, aes(x=PC1, y=PC2, col=batch )) + geom_point() + labs(x="PC1: 0.74", y="PC2: 0.09", title="Raw PAS qunatification data nuclear \n colored by batch ")
ggsave("../output/plots/QC_plots/NuclearPCA_colBatch.png",nucPCA_batch)

```

This shows that PC 2 is highly corrleated with batch,  

```{r}
nucPCA_mapped=ggplot(pca_nuc_df, aes(x=PC1, y=PC2, col=comb_mapped )) + geom_point() + labs(x="PC1: 0.74", y="PC2: 0.09", title="Raw PAS qunatification data nuclear \n colored by Mapped Read count")
ggsave("../output/plots/QC_plots/NuclearlPCA_colMapped.png",nucPCA_mapped)
```


##Q: Do the PAS read number recapitulate gene expression as it should?


Plot: scatter plot + fit (x-axis: gene TPM, y-axis: gene normalized PAS counts) total/nuclear separate

The TPM measurements come from the kalisto run I did on 18486. 

```{r}
tx2gene=read.table("../data/RNAkalisto/ncbiRefSeq.txn2gene.txt" ,header= F, sep="\t", stringsAsFactors = F)

txi.kallisto.tsv <- tximport("../data/RNAkalisto/abundance.tsv", type = "kallisto", tx2gene = tx2gene,countsFromAbundance="lengthScaledTPM" )
```

I need to get all of the peaks for 18486 and which gene they are in. Then I will take the gene average and divide by the number of mapped reads. 

```{r}
total_Cov_18486=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% filter(X18486_T>10) %>%  group_by(gene) %>% summarize(GeneSum=sum(X18486_T)) 

#%>% mutate(NormGenePeakCov=GeneSum/10819437)

```

Join with the transcript TPM  

```{r}
TXN_abund=as.data.frame(txi.kallisto.tsv$abundance) %>% rownames_to_column(var="gene")
colnames(TXN_abund)=c("gene", "TPM")

TXN_NormGene=TXN_abund %>% inner_join(total_Cov_18486,by="gene")
```

 
Plot distribution of each variable seperatly first to understand distribution:  
```{r}
summary(TXN_abund$TPM)
ggplot(TXN_abund, aes(x=TPM)) + geom_density(kernel="gaussian") + scale_x_log10()
```
```{r}
summary(total_Cov_18486$GeneSum)
ggplot(total_Cov_18486, aes(x=GeneSum)) + geom_density(kernel="gaussian")+ scale_x_log10()
```


Create a scatterplot:  
```{r}
corr_18486Tot=ggplot(TXN_NormGene, aes(x=log10(TPM+ .001), y= log10(GeneSum+.001))) + geom_point() + labs(title="Total", x="log10 RNA seq TPM", y="log10 Peak count sum per gene")+ geom_smooth(aes(x=log10(TPM +.001),y=log10(GeneSum+ .001)),method = "lm") + xlim(-2,6) + annotate("text",x=5, y=5,label="R2=.36")
       
corr_18486Tot       
summary(lm(log10(TPM +.001)~log10(GeneSum+ .001),TXN_NormGene)) 

```




Let me try this with the nuclear fraction:  

```{r}
nuclear_Cov_18486=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% filter(X18486_N>10) %>% group_by(gene) %>% summarize(GeneSum=sum(X18486_N)) 

#%>% mutate(NormGenePeakCov=GeneSum/11405271)

TXN_NormGene_Nuc=TXN_abund %>% inner_join(nuclear_Cov_18486,by="gene")



```

Create a scatterplot:  
```{r}


corr_18486Nuc=ggplot(TXN_NormGene_Nuc, aes(x=log(TPM+.001), y= log10(GeneSum+.001))) + geom_point() + geom_smooth(aes(x = log10(TPM +.001), y = log10(GeneSum+.001)),method = "lm",se=T) + labs(title=" Nuclear", x="log10 RNA seq TPM", y="log10 Peak Sum per gene") +xlim(-5,10) + annotate("text",x=-3, y=5,label="R2=.30")
corr_18486Nuc
summary(lm(log10(TPM +.001)~log10(GeneSum+ .001),TXN_NormGene_Nuc)) 
```


```{r}
title <- ggdraw() + draw_label("Correlation between TPM and 3' Seq \nNA18486", fontface='bold')

plots=plot_grid(corr_18486Tot,corr_18486Nuc)

CorrelationPlot18486=plot_grid(title,plots, ncol=1 , rel_heights = c(.1,1))
ggsave(file="../output/plots/QC_plots/CorrelationWKalisto18486.png",CorrelationPlot18486)
```



##Q: For each gene, what percentage of reads assigned fall within 1, 2, 3, etc… peaks, we would expect that for many genes >90% of the reads fall within 1 peak, for a few 2 peaks, etc…? 

Plot: Y-axis: Number of genes, X-axis: how many peaks is needed to “capture” 90%, 80%, … 50% of the reads assigned to that gene (using different colors). 

Start with analysis to see how many peaks are needed to capture 90% of the reads assigned to the gene. I will start by looking at the number of reads that map to peaks in genes. To do this I can group on genes in the peak coverage and get the sum.  


```{r}
nuclear_covBygene=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>% summarize(GeneSum=sum(X18486_N)) %>% mutate(per90=GeneSum*.9)%>% mutate(per80=GeneSum*.8)%>% mutate(per70=GeneSum*.7)%>% mutate(per60=GeneSum*.6)%>% mutate(per50=GeneSum*.5) 
total_covBygene=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>% summarize(GeneSum=sum(X18486_T))%>% mutate(per90=GeneSum*.9)%>% mutate(per80=GeneSum*.8)%>% mutate(per70=GeneSum*.7)%>% mutate(per60=GeneSum*.6)%>% mutate(per50=GeneSum*.5)
```


Write these out to use them in the script:  

```{r}
write.table(file="../data/UnderstandPeaksQC/Nuclear_PerCovbyGene.txt", nuclear_covBygene, quote=F, col.names = T, row.names = F)
write.table(file="../data/UnderstandPeaksQC/Total_PerCovbyGene.txt", total_covBygene, quote=F, col.names = T,row.names = F)
```



**Python method**  
I think the next step will be best accomplished with a python script where I can go line by line. I want to order the peak counts in a dictionary. This will  be double dictionary because I need gene and individual. I want the key to be the gene and the value to be an ordered list of the peak counts. I can ask how many elements it took to acieve the % read coverage.

1. Build dictionary

2. Refer to the *_covBygene gene number pairs to ask how many peaks get you to that number.  

3. Write out the gene and the number to capture the % of interest

4. The file should take a percentage between 50 and 90 and total/nuclear. I can then run the script on each value and join the dataframes in R. 

5. filter genes with 0 reads  


File is code/PeaksToCoverPerReads.py. I am stuck with the dictionary creation because the same values are there for each gene. All values are being added. The gene key is not being recognized.

**Try in R**
```{r}
#groupedNuclear=sumforeachgene %>% sort(ind) %>% cumulativesum %>% dividebygenesum %>% filter(only90) %>% count()
#remove genes with 0 count sum 
nuclear_90Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>%  arrange(gene,desc(X18486_N)) %>%  mutate(SUM = cumsum(X18486_N)) %>% full_join(nuclear_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.9) %>% tally() %>% rename("90"=n)

nuclear_80Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>%  arrange(gene,desc(X18486_N)) %>%  mutate(SUM = cumsum(X18486_N)) %>% full_join(nuclear_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.8) %>% tally() %>% rename("80"=n)

nuclear_70Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>%  arrange(gene,desc(X18486_N)) %>%  mutate(SUM = cumsum(X18486_N)) %>% full_join(nuclear_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.7) %>% tally() %>% rename("70"=n)

nuclear_60Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>%  arrange(gene,desc(X18486_N)) %>%  mutate(SUM = cumsum(X18486_N)) %>% full_join(nuclear_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.6) %>% tally() %>% rename("60"=n)

nuclear_50Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Nuclear_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_N) %>% group_by(gene) %>%  arrange(gene,desc(X18486_N)) %>%  mutate(SUM = cumsum(X18486_N)) %>% full_join(nuclear_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.5) %>% tally() %>% rename("50"=n)



```
Join these to plot them:  

```{r}
nuclear_PercentPeakCov= nuclear_90Cov %>% left_join(nuclear_80Cov, by="gene") %>% left_join(nuclear_70Cov, by="gene") %>% left_join(nuclear_60Cov, by="gene") %>% left_join(nuclear_50Cov, by="gene")

nuclear_PercentPeakCov_melt=melt(nuclear_PercentPeakCov,id.vars = "gene")
```

```{r}
nucPeakCov=ggplot(nuclear_PercentPeakCov_melt, aes(x=value,fill=variable))+ geom_histogram(position="dodge", bins=30) + labs(y="Number of Genes", x="Number of Peaks", title="Nuclear: Number of Peaks to capture % of Gene count") + facet_grid(~variable) + xlim(0,30)

ggplot(nuclear_PercentPeakCov_melt, aes(x=value,fill=variable, by=variable))+ geom_density(alpha=.4) + labs(y="Number of Genes", x="Number of Peaks", title="Nuclear: Number of Peaks to capture % of Gene count")

```

Try this with total:  

```{r}
total_90Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>%  arrange(gene,desc(X18486_T)) %>%  mutate(SUM = cumsum(X18486_T)) %>% full_join(total_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.9) %>% tally() %>% rename("90"=n)

total_80Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>%  arrange(gene,desc(X18486_T)) %>%  mutate(SUM = cumsum(X18486_T)) %>% full_join(total_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.8) %>% tally() %>% rename("80"=n)

total_70Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>%  arrange(gene,desc(X18486_T)) %>%  mutate(SUM = cumsum(X18486_T)) %>% full_join(total_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.7) %>% tally() %>% rename("70"=n)

total_60Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>%  arrange(gene,desc(X18486_T)) %>%  mutate(SUM = cumsum(X18486_T)) %>% full_join(total_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.6) %>% tally() %>% rename("60"=n)

total_50Cov=read.table("../data/PeakCounts/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.Total_fixed.fc", header=T, stringsAsFactors = F)[,1:7] %>% separate(Geneid, into=c("peak", "chr", "start", "end", "strand", "gene"), sep=":") %>% select(gene, X18486_T) %>% group_by(gene) %>%  arrange(gene,desc(X18486_T)) %>%  mutate(SUM = cumsum(X18486_T)) %>% full_join(total_covBygene,by="gene") %>% filter(GeneSum >0) %>% mutate(perSum=SUM/GeneSum) %>% mutate(perSum_lag=lag(perSum,1)) %>%  replace_na(list(perSum_lag =0)) %>% filter(perSum_lag<.5) %>% tally() %>% rename("50"=n)

```

Put together:  

```{r}
total_PercentPeakCov= total_90Cov %>% left_join(total_80Cov, by="gene") %>% left_join(total_70Cov, by="gene") %>% left_join(total_60Cov, by="gene") %>% left_join(total_50Cov, by="gene")

total_PercentPeakCov_melt=melt(total_PercentPeakCov,id.vars = "gene")
```

```{r}
totPeakCov=ggplot(total_PercentPeakCov_melt, aes(x=value,fill=variable))+ geom_histogram(position="dodge", bins=30) + labs(y="Number of Genes", x="Number of Peaks", title="Total: Number of Peaks to capture % of Gene count") + facet_grid(~variable) + xlim(0,30)


ggplot(total_PercentPeakCov_melt, aes(x=value,fill=variable, by=variable))+ geom_density(alpha=.4) + labs(y="Number of Genes", x="Number of Peaks", title="Nuclear: Number of Peaks to capture % of Gene count")

```


```{r}
PeakCovPerGeneCount=plot_grid(totPeakCov,nucPeakCov, ncol = 1)

ggsave(file="../output/plots/QC_plots/PeakCovPerGeneCount.png",PeakCovPerGeneCount)
```

##Q: What % of reads are assigned to a peak? Of these, what % of reads are assigned to a gene? 

Within 50bp of an exon (more relevant for total)?


I want to know the percent of reads that are assigned to our peaks. I can get this information from the peak feature counts summaries. In order to look at the reads assigned to genes I will need to use feature counts with the gene annotation file.  

* File: /project2/gilad/briana/threeprimeseq/data/filtPeakOppstrand_cov/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.fc.summary

Feature count takes in the bam files and an SAF annotation. For this one I used the peaks woth the transcript level annotation. I will fix the column names with python.  

fix_fc_summary.py
```{bash,eval=F}
infile= open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.fc.summary", "r")
fout = open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.fc.summary_fixed",'w')
for line, i in enumerate(infile):
    if line == 0:
        i_list=i.split()
        libraries=[i_list[0]]
        for sample in i_list[1:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        print(libraries)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n' )
    else:
        fout.write(i)
fout.close()
```


I care about Unassigned_NoFeatures and Assigned. These numbers add to the number of reads that map to the genome.   

```{r}
fc_peaks=read.table("../data/UnderstandPeaksQC/filtered_APApeaks_merged_allchrom_refseqGenes.Transcript_sm_quant.fc.summary_fixed", stringsAsFactors = F) %>% t()
fc_peaks=as.data.frame(fc_peaks)
colnames(fc_peaks)=as.character(unlist(fc_peaks[1,]))
fc_peaks=fc_peaks[-1,]
fc_peaks$Assigned=as.numeric(as.character(fc_peaks$Assigned))
fc_peaks$Unassigned_NoFeatures=as.numeric(as.character(fc_peaks$Unassigned_NoFeatures))
```

I need to separate the libraries by line and fraction.  
```{r}
fc_peaks=fc_peaks %>% separate(Status, into=c("line", "fraction"), sep="_") %>% mutate(PerReadPeak=Assigned/(Assigned+Unassigned_NoFeatures))
```


This number is the reads assigned to peaks out of all reads mapping to genome.  

I can now melt these data by line and fraction   

```{r}
fc_peaks_melt=melt(fc_peaks, id.vars = c("line", "fraction"))
fc_peaks_melt_PerRead=fc_peaks_melt %>% filter(variable=="PerReadPeak")
fc_peaks_melt_PerRead$value=as.numeric(fc_peaks_melt_PerRead$value)
```

```{r}
ggplot(fc_peaks_melt_PerRead,aes( x=line, y=value, by=fraction, fill=fraction))+ geom_col(pos="dodge") +theme(axis.text.x = element_text(angle = 90, hjust = 1),axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("deepskyblue3","darkviolet")) + labs(title="Percent of reads mapping to peaks by line and fraction", y="Reads mapping to peaks/all mapping reads")
```


It may be more interesting to look at this by fraction, with error bars.


```{r}
fc_peaks_melt_PerRead_byfrac= fc_peaks_melt_PerRead %>% group_by(fraction) %>% summarise(mean=mean(value), sd=sd(value))
```
Plot this:  
```{r}
ggplot(fc_peaks_melt_PerRead_byfrac,aes(x=fraction, y=mean, fill=fraction)) + geom_col()+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2)+ theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("deepskyblue3","darkviolet"))+ labs(title="Percent of reads mapping to peaks by fraction", y="Reads mapping to peaks/all mapping reads")
```

Now I want to look at how many reads map to gene. I will use the transcript annotations that I used for the peaks.  

* /project2/gilad/briana/genome_anotation_data/ncbiRefSeq_sm_noChr.sort.mRNA.bed 

I need to make this an SAF file.  
* GeneID
* Chr
* Start
* End 
* Strand  

RefSeqmRNA2SAF.py
```{bash,eval=F}
#python
from misc_helper import *
fout = file("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq_sm_noChr.sort.mRNA.SAF","w")
fout.write("GeneID\tChr\tStart\tEnd\tStrand\n")
for ln in open("/project2/gilad/briana/genome_anotation_data/ncbiRefSeq_sm_noChr.sort.mRNA.bed"):
    chrom, start, end, gene, score, strand = ln.split()
    start_i=int(start)
    end_i=int(end)
    fout.write("%s\t%s\t%d\t%d\t%s\n"%(gene, chrom, start_i, end_i, strand))
fout.close()
```


ref_geneTranscript_fc.sh   
```{bash,eval=F}
#!/bin/bash

#SBATCH --job-name=ref_geneTranscript_fc
#SBATCH --account=pi-yangili1
#SBATCH --time=24:00:00
#SBATCH --output=ref_geneTranscript_fc.out
#SBATCH --error=ref_geneTranscript_fc.err
#SBATCH --partition=broadwl
#SBATCH --mem=12G
#SBATCH --mail-type=END


module load Anaconda3
source activate three-prime-env

featureCounts -O -a /project2/gilad/briana/genome_anotation_data/ncbiRefSeq_sm_noChr.sort.mRNA.SAF -F SAF -o /project2/gilad/briana/threeprimeseq/data/UnderstandPeaksQC/RefSeqTranscript_AllLibraries.fc /project2/gilad/briana/threeprimeseq/data/sort/*sort.bam -s 2
```



fix_Genefc_summary.py
```{bash,eval=F}
infile= open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/UnderstandPeaksQC/RefSeqTranscript_AllLibraries.fc.summary", "r")
fout = open("/Users/bmittleman1/Documents/Gilad_lab/threeprimeseq/data/UnderstandPeaksQC/RefSeqTranscript_AllLibraries.fc.summary_fixed",'w')
for line, i in enumerate(infile):
    if line == 0:
        i_list=i.split()
        libraries=[i_list[0]]
        for sample in i_list[1:]:
            full = sample.split("/")[7]
            samp= full.split("-")[2:4]
            lim="_"
            samp_st=lim.join(samp)
            libraries.append(samp_st)
        print(libraries)
        first_line= "\t".join(libraries)
        fout.write(first_line + '\n' )
    else:
        fout.write(i)
fout.close()
```


```{r}
fc_gene_peaks=read.table("../data/UnderstandPeaksQC/RefSeqTranscript_AllLibraries.fc.summary_fixed", stringsAsFactors = F) %>% t()
fc_gene_peaks=as.data.frame(fc_gene_peaks)
colnames(fc_gene_peaks)=as.character(unlist(fc_gene_peaks[1,]))
fc_gene_peaks=fc_gene_peaks[-1,]
fc_gene_peaks$Assigned=as.numeric(as.character(fc_gene_peaks$Assigned))
fc_gene_peaks$Unassigned_NoFeatures=as.numeric(as.character(fc_gene_peaks$Unassigned_NoFeatures))

```

I need to separate the libraries by line and fraction.  
```{r}
fc_gene_peaks=fc_gene_peaks %>% separate(Status, into=c("line", "fraction"), sep="_") %>% mutate(PerReadPeak=Assigned/(Assigned+Unassigned_NoFeatures))
```

Melt this: 


```{r}
fc_gene_peaks_melt=melt(fc_gene_peaks, id.vars = c("line", "fraction"))
fc_gene_peaks_PerRead=fc_gene_peaks_melt %>% filter(variable=="PerReadPeak")
fc_gene_peaks_PerRead$value=as.numeric(fc_gene_peaks_PerRead$value)
```


GGplot:  

```{r}
ggplot(fc_gene_peaks_PerRead,aes( x=line, y=value, by=fraction, fill=fraction))+ geom_col(pos="dodge") +theme(axis.text.x = element_text(angle = 90, hjust = 1),axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("deepskyblue3","darkviolet")) + labs(title="Percent of reads mapping to Transcripts by line and fraction", y="Reads mapping to transcripts/all mapping reads")
```

Do this by fraction.   

```{r}
fc_gene_peaks_PerRead_byfrac= fc_gene_peaks_PerRead %>% group_by(fraction) %>% summarise(mean=mean(value), sd=sd(value))
```
Plot this:  
```{r}
ggplot(fc_gene_peaks_PerRead_byfrac,aes(x=fraction, y=mean, fill=fraction)) + geom_col()+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2)+ theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("deepskyblue3","darkviolet"))+ labs(title="Percent of reads mapping to Transcripts by fraction", y="Reads mapping to Transcripts/all mapping reads")
```


It would be nice to have this in one plot. In order to do this I want to join the PerReadPeak from both and melt. this way the variable can be peak or transcript.  

```{r}
fc_peaks_sel=fc_peaks %>% select(c("line", "fraction", "PerReadPeak"))

fc_gene_peaks_sel=fc_gene_peaks %>% select(c("line", "fraction", "PerReadPeak"))

fcGene_and_Transcript=fc_peaks_sel %>% left_join(fc_gene_peaks_sel, by=c("line","fraction"))

colnames(fcGene_and_Transcript)=c("Line", "Fraction", "Peaks", "Genes")


fcGene_and_Transcript_melt=melt(fcGene_and_Transcript, id.vars=c("Line","Fraction"))


fcGene_and_Transcript_melt_sum=fcGene_and_Transcript_melt %>% group_by(Fraction,variable) %>% summarise(mean=mean(value), sd=sd(value))
```


```{r}

reads2featuresPlot=ggplot(fcGene_and_Transcript_melt_sum,aes(x=Fraction, y=mean, fill=Fraction)) + geom_col()+ geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2)+ theme(axis.text.y = element_text(size=12),axis.title.y=element_text(size=10,face="bold"), axis.title.x=element_text(size=12,face="bold"))+ scale_fill_manual(values=c("deepskyblue3","darkviolet"))+ labs(title="Percent of reads mapping to feature by fraction", y="Reads mapping to Feature/all mapping reads") + facet_grid(~variable)

reads2featuresPlot

ggsave(file="../output/plots/QC_plots/reads2featuresPlot.png", reads2featuresPlot)
```

